% Text Analysis
% Laurence A. F. Park
% March 22, 2005

<!-- Setting up R -->
`ro warning=FALSE, dev="png", fig.cap="", cache=FALSE or`






# What is text

## This is text

Unstructured Data

### This is a block!

The contents of the block

#### What is this?

> This text should
> appear in a block.

## Storing Text

  \begin{columns}
    \begin{column}{.5\linewidth}
      \begin{block}{Some equation}
         \begin{align*}
           a &= b\\
             &= d\\
             &= f
         \end{align*}
      \end{block}
    \end{column}
    \begin{column}{.5\linewidth}
      \begin{tabular}{cc}\hline
        some text & some text \\\hline
      \end{tabular}
    \end{column}
  \end{columns}

## Searching through text

<!-- read external r code -->
```{r reading, echo=FALSE}
read_chunk("script.r")
```

## The first part of my R script
Here I can generate my data
```{r}
<<gen-dat>>
```

## Results
An now the reults are plotted
```{r plot-fig, result="asis"}
<<plot>>
```


## Term Frequency Weighting


# In the morning

## Getting up

- Turn off alarm
- Get out of bed

## Breakfast

- Eat eggs
- Drink coffee

# In the evening

## Dinner

- Eat spaghetti
- Drink wine

------------------


## Going to sleep

- Get in bed
- Count sheep
