% Revision
% Laurence A. F. Park
% 20th of September 2013

```{r setup, include=FALSE}
opts_chunk$set(fig.path='figure/summary-summary-')
opts_chunk$set(dev = 'pdf', fig.cap="")
```


# Updated exam questions

### Changes to the Unit


Over the years we have made changes to the unit, so some of the
previous years exam questions may not be directly relevant to us.

We will go a few previous exam questions and modify them to made them
appropriate for this year.


### Naive Bayes for Sentiment Analysis

There was a previous exam question (2013 Q8) focused on Sentiment
Analysis using Naive Bayes classification. Since the time of that
exam, we have simplified the Naive Bayes process, so the existing
solution is not what we have taught. We will now show to answer this
question using the method taught this semester.

### 2013 Question 8

The Funfun toothpaste company wants to gauge the sentiment towards
their toothpaste.  Given the following set of labelled tweets:

\vspace{1em}
\begin{columns}[t]
\begin{column}{6cm}
Positive tweets:
\begin{itemize}
\item My teeth shine \#funfun
\item \#funfun love my fun teeth
\item \#funfun is fun fun
\end{itemize}
\end{column}
\begin{column}{6cm}
Negative Tweets:
\begin{itemize}
\item No shine \#funfun
\item No love fun fun
\item Where is my teeth shine \#funfun
\end{itemize}
\end{column}
\end{columns}

\begin{enumerate} 
\item Calculate and tabulate the word distribution for both positive and negative sentiment ($P(\text{term}\vert\text{sentiment})$).
\item Calcuate the probability of the tweet ``fun teeth shine'' ($P(\text{tweet})$).
\item Compute the log likelihood ratio of the tweet ``fun teeth shine'' being positive versus it being negative, using Naive Bayes classification ($\log(P(\text{sentiment} = \text{Pos}\vert \text{tweet})/P(\text{sentiment} = \text{Neg}\vert \text{tweet}))$).
\item Determine if the tweet ``fun teeth shine'' has positive or negative sentiment, based on the log likelihood ratio.
\end{enumerate}

### 2013 Question 8 Answers

- $P(\text{term}\vert\text{sentiment}) =$

\begin{center}
\begin{tabular}{rccccccccc}
\toprule
    & \#funfun & fun & is  & love & my  & no  & shine & teeth & where \\
\midrule
Pos & 1        & 2/3 & 1/3 & 1/3  & 2/3 & 0   & 1/3   & 2/3   & 0     \\
Neg & 2/3      & 1/3 & 1/3 & 1/3  & 1/3 & 2/3 & 2/3   & 1/3   & 1/3   \\
\bottomrule
\end{tabular}
\end{center}

```{r echo = FALSE}
pSentPos = 0.5
pSentNeg = 0.5

# fun teeth shine
pTweetP = 2/3*2/3*1/3
pTweetN = 1/3*1/3*2/3

# fun teeth shine
#pTweet = (2/3*1/2 + 1/3*1/2)*(2/3*1/2+ 1/3*1/2)*(2/3*1/2 + 1/3*1/2)
pTweet = pTweetP * pSentPos + pTweetN * pSentNeg


pPTweet = pTweetP*pSentPos/pTweet
pNTweet = pTweetN*pSentNeg/pTweet

s = log(pPTweet/pNTweet)
```

- $P(\text{tweet}) = `r pTweet`$
- $\log\left (\frac{P(\text{sentiment} = \text{Pos}\vert \text{tweet})}{P(\text{sentiment} = \text{Neg}\vert \text{tweet})}\right ) = `r s`$

\pause
Let's show how to get these answers.

### Randomisation Distributions

This sememester, we introduced hypothesis testing using randomisation
distributions. So there are no previous exam questions that use
randomisation distributions.

Hypothesis testing using randomisation distributions appear in:

\begin{center}
\begin{tabular}{clc}
\toprule
Week & Test & Statistic \\
\midrule
3    & Exposure analysis: test for independence & $\chi^2$ \\
10   & Trends test for independence & $\chi^2$ \\
11   & Impact difference in means & $\mu_{\text{After}} - \mu_{\text{Before}}$ \\
11   & Impact test for interaction & $\gamma$ \\
\bottomrule
\end{tabular}
\end{center}

We can't ask you to compute a randomisation distribution without
access to a computer. But we can ask you to describe the process and
to analyse the results.


### Example Randomisation Question

```{r echo=FALSE}
set.seed(56223)
n = 5
before = rpois(n, 200)
after = rpois(n, 100)
cbefore = rpois(n, 100) 
cafter = rpois(n, 80)
suppressWarnings(library(dae, quietly=TRUE))
xs = c(cbefore, cafter, before, after)
when = factor(c(rep("before", n), rep("after", n), rep("before", n),
       rep("after", n)), levels = c("before", "after"))
company = factor(c(rep("control", 2 * n), rep("impact", 2 * n)),
	levels = c("control","impact"))

m = aov(xs ~ when + company)
f = predict(m)
r = residuals(m)
### permute residuals and replace
R = 1000
gSet = rep(0,R)
for (m in 1:R) {
    y = f + sample(r)
    gSet[m] = yates.effects(aov(y ~ when * company))[3]
}

gVal = yates.effects(aov(xs ~ when * company))[3]
pVal = mean(abs(gSet) > abs(gVal))

```
#### Question

The number of tweets containing "#jumpy" were counted for five days
before and after the Jumpy advertising campaign. A set of control
tweets were also counted. The counts are listed below. We want to
determine if the advertising campaign had an effect of the number of
tweets containing "#jumpy".


\begin{center}
\begin{tabular}{lcc}
\toprule
& Control & Impact \\
\midrule
Before & `r paste(cbefore, collapse=", ")` & `r paste(before, collapse=", ")`  \\
After  & `r paste(cafter, collapse=", ")` & `r paste(after, collapse=", ")`  \\
\bottomrule
\end{tabular}
\end{center}


1. State the Null and alternative hypothesis of the test that needs to
   be conducted.
2. State which test statistic should be used for the test and provide
   its equation.


### Example Randomisation Question

#### Question

3. Describe the randomisation process to obtain the distribution of
   the test statistic.
4. Given the following randomisation distribution, what is the
   conclusion of the test?

```{r echo=FALSE, fig.width=4.2, fig.height=2.0, fig.cap="Distribution of sample interaction."}
qplot(gSet,geom="histogram",binwidth=1) + theme_grey(base_size = 10) + theme(legend.position = "none") + xlab("Randomisation statistic")
```

### Document Models

Document models are covered in the week 4 lecture.
In previous years we taught the document models:

- Vector Space Model: TF-IDF
- Divergence from Randomness: Binomial Distribution

This year we updated the document models to:

- Vector Space Model: TF-IDF
- Language Model using Dirichlet Smoothing

If you see previous exam questions with "Divergence from Randomness"
then adjust them to another model that we have learnt.


# Revision

### Revision Questions

For the remainder of this lecture, we will work though a set of past
exam questions.

- Text Indexing: 2015, Q2: change "the Divergence from Randomness weight (using the
  Binomial Distribution)" to "TF-IDF weight"
- Graphs: 2014, Q1.
- Visualisation: 2013, Q3.
- Clustering: 2015, Q4.
- Link Analysis: 2013, Q5.
- Trends: 2014, Q8.
- Sentiment: 2015, Q8.

We may not get through all of these questions. If we don't, all of the
solutions are available in vUWS and questions can be asked using the
Unit email mailing list.



# Your Future

### Where to go next

If you are in your final year, you should have a good idea of what you want to
do next year.

The most likely choices are:

- Get a job
- Enrol in a higher degree.

### Industry

There are many business within Australia and overseas that have large
amounts of data, but do not know how to analyse it (or are using very
simple analysis methods) to improve their business.

- Banks
- Supermarkets
- Insurance Companies
- Telecommunication Companies

Businesses with large customer databases are crying out for people with data
mining skills.


### Honours and PhD

Honours and PhD degrees are research degrees. Research degrees have no
lectures, but require the student to investigate and research a new
topic in the field of choice. Research commonly leads to:

- new discoveries, 
- publishing the discoveries in international journals, and
- being sent overseas to present the new discoveries to an international audience.

After you finish a PhD, you get to be called Doctor!

Social Web Analysis is a thriving research topic in the domains of
Data Mining and Machine Learning.

If you are interested in further study in this field (or similar fields) speak to Laurence and Glenn.

Current research can be found using Google Scholar:
<http://scholar.google.com.au/>


### Data Science Degree

If this unit has interested you, and you are unsure if you want to go
on to do research, you can extend your degree by \alert{one year} and
obtain a Bachelor of Data Science along with your existing degree.


If you are interested in completing a Bachelor of Data Science, speak
to Laurence ot Glenn.

Note that you must enrol in the additional degree \alert{before} you
finish your existing degree. The Bachelor of Data Science cannot be
completed alone.  So if this is your final year, you must enrol in the
Bachelor of Data Science before you finish this semester.



# Exam details

### Date and Venue

The date and venue of the final exam are set by University Timetabling.

Check the exam timetable Web site to find when and where the exam will be held.

Don't rely on a friend giving you this information, look it up yourself!

For more information: <http://uws.edu.au/uws_students>

### Open Book Exam

The exam is open book, meaning that you can bring material into the exam room.

However, be sensible. The more paper you bring in, the harder it is to find a piece of information.

Note: no phones, computers or iPads are allowed in the exam room.

### Duration and Questions

The exam will last two hours.

It contains \alert{seven} questions, each worth eight marks.

When reading through the exam, each question will be labelled with the
number of marks allocated to it.  If there are sub questions, the
break down of the marks will be shown.

Two hours/seven questions $\approx 17$ minutes per question.

### Sample Exam

A sample exam and previous years exams are available in vUWS. There have
been changes to the Unit content over the years, so some questions may
not be relevant for us. The list of relevant questions is in the Exams
section of vUWS.

The solutions to the sample exam and previous years exams are
available. The solutions to last years exam will be available next
week.

Don't waste these practice exams!


### Thanks

Thank-you to all students. We had fun setting up and teaching this unit, we
hope you all enjoyed learning about Social Web Analytics.

We hope we don't see you next year!
