% Introduction to Graphs (Definition, Graph statistics, Storage)
% Laurence A. F. Park
% March 22, 2005

```{r setup, include=FALSE}
opts_chunk$set(dev = 'pdf', fig.cap="")
```


## Social Networks

It is the *Network* that makes Social Networking sociable. Without the
social network, the Web would be less dynamic. There would be less
information flow, meaning that the Web would be less useful.

A network is also called a graph. In this lecture we will examine
basic graph theory, to allow us to analyse the social network.


# What is a graph?


## Not a graph

This was a graph when you were in high school.
This is now called a *plot*.

\vspace{-6em}
```{r, echo=FALSE, fig.width=5, fig.height=4}
x = (1:10)
y = x^3 %% 10
#par(mar = (c(3.5, 3.5, 4, 2) + 0.1))
par(mgp = c(1.5, 0.4, 0))
plot(x, y, xlab="Some numbers", ylab="Some other numbers", type="b", cex.lab=0.8, cex.axis=0.8,)
```

## A graph

Now that you have passed high school, we can let you know that this is a graph.

\vspace{-2em}
```{r, echo=FALSE, fig.width=5, fig.height=2.8}
library("igraph")
par(mar = c(0,0,0,0))
g = erdos.renyi.game(10,p=0.4)
V(g)$label.cex = 0.8
plot(g, layout=layout.fruchterman.reingold, vertex.size = 15)
```

## Undirected Graph

An undirected graph has no direction on the edges.

\vspace{-2em}
```{r, echo=FALSE, fig.width=5, fig.height=2.8}
par(mar = c(0,0,0,0))
g = erdos.renyi.game(10,p=0.4)
V(g)$label.cex = 0.8
plot(g, layout=layout.fruchterman.reingold, vertex.size = 15)
```

## Directed Graph

An Directed graph has direction on the edges.

\vspace{-2em}
```{r, echo=FALSE, fig.width=5, fig.height=2.8}
par(mar = c(0,0,0,0))
g = erdos.renyi.game(10,p=0.4, directed=TRUE)
V(g)$label.cex = 0.8
plot(g, layout=layout.fruchterman.reingold, vertex.size = 15)
```

## Weighted Graph

A Weighted Graph has weighted edges.

\vspace{-2em}
```{r, echo=FALSE, fig.width=5, fig.height=2.8}
par(mar = c(0,0,0,0))
g = erdos.renyi.game(10,p=0.4, directed=FALSE)
V(g)$label.cex = 0.8
E(g)$weight <- rpois(ecount(g),2) + 1
plot(g, layout=layout.fruchterman.reingold, vertex.size = 15,edge.label=E(g)$weight)
```




## Formal Definition

A graph is a set of vertices, connected by edges.
A graph is given by $G = (V,E)$, where $V$ is the set of vertices:
$$
	V = \{v_1, v_2, \ldots, v_n \}
$$
and $E$ is the set of edges (pairs of vertices):
$$
	E= \{e_1, e_2, \ldots, e_n\}, \quad e_i = \{v_i, v_j\}, \quad v_i, v_j \in V
$$


\begin{columns}
\begin{column}{0.45\textwidth}
We define this graph as:
\begin{itemize}
\item $V = \{v_1, v_2, v_3, v_4\}$
\item $E = \{\{v_1, v_2\}, \{v_1, v_3\}, \{v_1, v_4\}, \{v_2, v_4\} \}$
\end{itemize}
The number of vertices is $\lvert V\rvert$ (the size of set $V$).
The number of edges is $\lvert E\rvert$ (the size of set $E$).
\end{column}
\begin{column}{0.45\textwidth}
%```{r, echo=FALSE, fig.width=1, fig.height=1}
%g = graph.formula(A-B, A-C, A-D, B-D)
%pdf("simple_graph.pdf", width=3, height=3)
%par(mar = c(0,0,0,0))
%V(g)$label.cex = 2
%plot(g, vertex.label = c(expression("v"[1]),expression("v"[2]),expression("v"[3]),expression("v"[4])), 
%layout=layout.fruchterman.reingold, vertex.size = 50)
%dev.off()
%```
\includegraphics[scale=0.6]{simple_graph.pdf}
\end{column}
\end{columns}

## Representing Graphs using an Adjacency List

A graph is a set of vertices and edges. To store a graph, we do not
need to store the image, we only need to store the vertices and edges.
A simple method of representing a graph is an Adjacency List, listing
each pair of vertices connected with edges.

### Example Adjacency List
\begin{columns}
\begin{column}{0.45\textwidth}
We define this graph as:
\begin{itemize}
\item $v_1$, $v_2$
\item $v_1$, $v_3$
\item $v_1$, $v_4$
\item $v_2$, $v_4$
\end{itemize}
\end{column}
\begin{column}{0.45\textwidth}
\includegraphics[scale=0.6]{simple_graph.pdf}
\end{column}
\end{columns}


## Representing Graphs using an Adjacency Matrix

An Adjacency Matrix $A$ is a $|V|\times|V|$ matrix (where $|V|$ is the
number of vertices). The element $a_{i,j}$ in the $i$ row and $j$th
column is 1 if there is an edge between the $i$the and $j$th vertice
and 0 if there is no edge.

### Example Adjacency Matrix
\begin{columns}
\begin{column}{0.45\textwidth}
We define this graph as:
\begin{align*}
A = \left [
\begin{array}{cccc}
	0 & 1 & 1 & 1 \\
	1 & 0 & 0 & 1 \\
	1 & 0 & 0 & 0 \\
	1 & 1 & 0 & 0 \\
\end{array}
\right ]
\end{align*}
\end{column}
\begin{column}{0.45\textwidth}
\includegraphics[scale=0.6]{simple_graph.pdf}
\end{column}
\end{columns}

## Adjacency Matrix Problem

### Problem

Write down the adjacency matrix for the following graph:

```{r, echo=FALSE, fig.width=1.5, fig.height=1.5}
g = erdos.renyi.game(5,p=0.7, directed=FALSE)
par(mar = c(0,0,0,0))
V(g)$label.cex = 0.8
plot(g, vertex.label = c(expression("v"[1]),expression("v"[2]),expression("v"[3]),expression("v"[4]),
expression("v"[5])),
layout=layout.fruchterman.reingold, vertex.size = 35)
```



# Graph Structure

## Subgraphs

A subgraph is a subset of the vertices of a graph and all interconnected edges.

### All Subgraphs

There are 16 different subgraphs of the following graph. List all 16.

```{r, echo=FALSE, fig.width=1.5, fig.height=1.5}
g = graph.formula(A-B, A-C, A-D, B-D)
par(mar = c(0,0,0,0))
V(g)$label.cex = 0.8
plot(g, vertex.label = c(expression("v"[1]),expression("v"[2]),expression("v"[3]),expression("v"[4])), 
layout=layout.fruchterman.reingold, vertex.size = 35)
```

## Complete Graph (Clique)

A set of vertices where all vertices have edges between them, is
called a complete graph, or a clique.


```{r, echo=FALSE, fig.width=1.5, fig.height=1.5}
g = graph.formula(A-B, A-C, A-D, B-C, B-D, C-D)
par(mar = c(0,0,0,0))
V(g)$label.cex = 0.8
plot(g, vertex.label = c(expression("v"[1]),expression("v"[2]),expression("v"[3]),expression("v"[4])), 
layout=layout.fruchterman.reingold, vertex.size = 35)
```

## Egocentric Graphs

An egocentric graph is a subgraph that is centred around a chosen vertex.

A degree 1 egocentric graph is a subgraph that contains the chosen
vertex, all connected vertices and the edges between them.

### Degree 1 egocentric graph of $v_1$

```{r, echo=FALSE, fig.width=1.5, fig.height=1.5}
g = graph.formula(A-B, A-C, A-D, B-E, B-D, C-D, C-E, C-F, D-F, F-G)
E(g)$color = "#AAAAAA"
V(g)$frame.color = "#AAAAAA"
V(g)$color = "#DDDDDD"

E(g)[c(1,2,3)]$color = "#000000"
V(g)[c("A","B","C","D")]$color = "#00BBBB"
V(g)[c("A","B","C","D")]$frame.color = "#000000"

par(mar = c(0,0,0,0))
V(g)$label.cex = 0.8
plot(g, vertex.label = c(expression("v"[1]),expression("v"[2]),expression("v"[3]),expression("v"[4]),
expression("v"[5]),expression("v"[6]),expression("v"[7])),
layout=layout.fruchterman.reingold, vertex.size = 35)
```

## Egocentric Graphs

A degree 1.5 egocentric graph is a subgraph that contains the chosen
vertex, all connected vertices and the edges between all vertices.

### Degree 1.5 egocentric graph of $v_1$

```{r, echo=FALSE, fig.width=1.5, fig.height=1.5}
g = graph.formula(A-B, A-C, A-D, B-E, B-D, C-D, C-E, C-F, D-F, F-G)
E(g)$color = "#AAAAAA"
V(g)$frame.color = "#AAAAAA"
V(g)$color = "#DDDDDD"

E(g)[c(1,2,3,4,6)]$color = "#000000"
V(g)[c("A","B","C","D")]$color = "#00BBBB"
V(g)[c("A","B","C","D")]$frame.color = "#000000"

par(mar = c(0,0,0,0))
V(g)$label.cex = 0.8
plot(g, vertex.label = c(expression("v"[1]),expression("v"[2]),expression("v"[3]),expression("v"[4]),
expression("v"[5]),expression("v"[6]),expression("v"[7])),
layout=layout.fruchterman.reingold, vertex.size = 35)
```

## Egocentric Graphs

A degree 2 egocentric graph is a subgraph that contains the chosen
vertex, all connected vertices and edges that are at most two hops
from the chosen vertex.

### Degree 2 egocentric graph of $v_1$

```{r, echo=FALSE, fig.width=1.5, fig.height=1.5}
g = graph.formula(A-B, A-C, A-D, B-E, B-D, C-D, C-E, C-F, D-F, F-G)
E(g)$color = "#AAAAAA"
V(g)$frame.color = "#AAAAAA"
V(g)$color = "#DDDDDD"

E(g)[c(1,2,3,4,5,6,7,8,9)]$color = "#000000"
V(g)[c("A","B","C","D","E","F")]$color = "#00BBBB"
V(g)[c("A","B","C","D","E","F")]$frame.color = "#000000"

par(mar = c(0,0,0,0))
V(g)$label.cex = 0.8
plot(g, vertex.label = c(expression("v"[1]),expression("v"[2]),expression("v"[3]),expression("v"[4]),
expression("v"[5]),expression("v"[6]),expression("v"[7])),
layout=layout.fruchterman.reingold, vertex.size = 35)
```

## Extract the Graph

### Find the degree 1 and 1.5 egocentric graph centred at vertex $1$.

```{r, echo=FALSE, fig.width=5, fig.height=2.8}
par(mar = c(0,0,0,0))
g = erdos.renyi.game(10,p=0.4)
V(g)$label.cex = 0.8
plot(g, layout=layout.fruchterman.reingold, vertex.size = 15)
```


# Graph Parameters


## Vertex Degree

The degree of a vertex is measured as the number of edges that connect to the vertex.

```{r, echo=FALSE, fig.width=1.5, fig.height=1.5}
g = graph.formula(A-B, A-C, A-D, B-E, B-D, C-D, C-E, C-F, D-F, F-G)

par(mar = c(0,0,0,0))
V(g)$label.cex = 0.8
plot(g, vertex.label = c(expression("v"[1]),expression("v"[2]),expression("v"[3]),expression("v"[4]),
expression("v"[5]),expression("v"[6]),expression("v"[7])),
layout=layout.fruchterman.reingold, vertex.size = 35)
```
The degree of $v_1$ is 3. The degree of $v_7$ is 1.

### Highest degree

Which vertex has the largest degree?

## Degree Distribution

By examining the degree of all of the vertices, we can compute the
degree distribution of the graph.

```{r, echo=FALSE, fig.width=1.5, fig.height=1.5}
g = graph.formula(A-B, A-C, A-D, B-E, B-D, C-D, C-E, C-F, D-F, F-G)

par(mar = c(0,0,0,0))
V(g)$label.cex = 0.8
plot(g, vertex.label = c(expression("v"[1]),expression("v"[2]),expression("v"[3]),expression("v"[4]),
expression("v"[5]),expression("v"[6]),expression("v"[7])),
layout=layout.fruchterman.reingold, vertex.size = 35)
```

\begin{center}
\begin{tabular}{lcccccc}
\toprule
Degree & 0 & 1 & 2 & 3 & 4 & 5 \\
Frequency & \\
\bottomrule
\end{tabular}
\end{center}

## Graph Density

A dense graph has many edges. A sparse graph has only a few edges. The
density of a graph is the ratio of the edges that exist, compared to
the possible edges.

### Possible number of edges

Given $\lvert V\rvert$ vertices, the maximum number of edges is:
$$
	{\sum_{i = 1}^{\lvert V\rvert} i} = \lvert V\rvert (\lvert V\rvert - 1)/2
$$

### Graph Density

The density $D$ of graph $G = (V,E)$ is:
$$
	D = \frac{\text{Number of edges}}{\text{Possible number of edges}} = \frac{2\lvert E\rvert}{\lvert V\rvert (\lvert V\rvert - 1)}
$$ 


## Path

A path is a sequence of edges that are connected with vertices.  Paths
can contain the same edge multiple times. Therefore, we can make a
path from a given start vertex and end vertex as long as we like.

### Two paths from $v_1$ to $v_6$
```{r, echo=FALSE, fig.width=1.5, fig.height=1.5}
g = graph.formula(A-B, A-C, A-D, B-E, B-D, C-D, C-E, C-F, D-F, F-G)
E(g)$color = "#AAAAAA"

E(g)[c(3,9)]$color = "#AA0000"
#E(g)[c(2,8)]$color = "#00AA00"
E(g)[c(1,5,7,8)]$color = "#0000AA"

par(mar = c(0,0,0,0))
V(g)$label.cex = 0.8
plot(g, vertex.label = c(expression("v"[1]),expression("v"[2]),expression("v"[3]),expression("v"[4]),
expression("v"[5]),expression("v"[6]),expression("v"[7])),
layout=layout.fruchterman.reingold, vertex.size = 35)
```

## Shortest Path

The shortest path between two vertices is the one that contains the
least number of edges. Note the the shortest path may not be unique
(there may be multiple paths that are just as short).

### Two shortest paths from $v_1$ to $v_6$
```{r, echo=FALSE, fig.width=1.5, fig.height=1.5}
g = graph.formula(A-B, A-C, A-D, B-E, B-D, C-D, C-E, C-F, D-F, F-G)
E(g)$color = "#AAAAAA"

E(g)[c(3,9)]$color = "#AA0000"
E(g)[c(2,8)]$color = "#0000AA"


par(mar = c(0,0,0,0))
V(g)$label.cex = 0.8
plot(g, vertex.label = c(expression("v"[1]),expression("v"[2]),expression("v"[3]),expression("v"[4]),
expression("v"[5]),expression("v"[6]),expression("v"[7])),
layout=layout.fruchterman.reingold, vertex.size = 35)
```

## Graph Diameter

The diameter of a circle is the longest straight line that can fit in
the circle.

The diameter of a graph is the longest shortest path in the graph.

To find the diameter, we must find the shortest paths between all
pairs of vertices. The diameter is the length of the longest of these
paths.

### All shortest paths

\begin{columns}
\begin{column}{0.45\textwidth}
\begin{itemize}
\item $v_1$ to $v_2$: 1
\item $v_1$ to $v_3$: 1
\item $v_1$ to $v_4$: 1
\item $v_2$ to $v_3$: 2
\item $v_2$ to $v_4$: 1
\item $v_3$ to $v_4$: 2
\end{itemize}
The diameter of the graph is 2.
\end{column}
\begin{column}{0.45\textwidth}
\includegraphics[scale=0.6]{simple_graph.pdf}
\end{column}
\end{columns}


## Graph Parameter Problem

### Problem

Compute the density and diameter of the following graph.

```{r, echo=FALSE, fig.width=1.5, fig.height=1.5}
g = graph.formula(A-B, A-C, A-D, B-D, B-E, D-E, B-F)
par(mar = c(0,0,0,0))
V(g)$label.cex = 0.8
plot(g, vertex.label = c(expression("v"[1]),expression("v"[2]),expression("v"[3]),expression("v"[4]),
expression("v"[5]),expression("v"[6])),
layout=layout.fruchterman.reingold, vertex.size = 35)
```


# Centrality


## Measures of Centrality

A measure of centrality directs us to the centre of an object,
therefore the location of the centre depends on how we define the
centre.

For continuous numerical data, we can compute the centre as the mean,
median or mode, depending on how we define the centre. For graphs there are
also many measures of centrality.

There is usually no clear centre of a graph, therefore centrality methods
provide a centrality score to all vertices, where the greater the score,
the more likely the centre exists at that associated vertex.

## Degree Centrality

The simplest measure of centrality is *Degree Centrality*. The
centrality score for each vertex is equal to the number of edges
connected to it. The reasoning for this measure is, the greater the
degree of a vertex, the more likely that information will flow through
it.
$$
	C_{\text{degree}}(v) = \text{deg}(v)
$$

### Example

\begin{columns}
\begin{column}{0.45\textwidth}
\begin{itemize}
\item $\text{deg}(v_1) = 3$
\item $\text{deg}(v_2) = 2$
\item $\text{deg}(v_3) = 1$
\item $\text{deg}(v_4) = 2$
\end{itemize}
The centre of the graph is $v_1$.
\end{column}
\begin{column}{0.45\textwidth}
\includegraphics[scale=0.5]{simple_graph.pdf}
\end{column}
\end{columns}

## Degree Centrality Problem

### Problem

Find the most central vertex, using degree centrality.

```{r, echo=FALSE, fig.width=1.5, fig.height=1.5}
#g = graph.formula(A-B, A-C, A-D, B-D)
g = erdos.renyi.game(7,p=0.6, directed=FALSE)
par(mar = c(0,0,0,0))
V(g)$label.cex = 0.8
plot(g, vertex.label = c(expression("v"[1]),expression("v"[2]),expression("v"[3]),expression("v"[4]),
expression("v"[5]),expression("v"[6]),expression("v"[7])),
layout=layout.fruchterman.reingold, vertex.size = 35)
```



## Closeness Centrality

A vertex should be considered the centre if we can reach all other
vertices from it with the shortest number of hops.

If we define $d(v_i, v_j)$ as the number of hops along the shortest
path from $v_i$ to $v_j$, then we define closeness centrality as:
$$
	C_{\text{closeness}}(v) = {\sum_{u \in V} d(v,u)}
$$

### Example

\begin{columns}
\begin{column}{0.45\textwidth}
\begin{itemize}
%\item $d(v_1,v_2) + d(v_1,v_3) + d(v_1,v_4) = 1 + 1 + 1 = 3$
\item $C_{\text{closeness}}(v_1) = 1 + 1 + 1 = 3$
\item $C_{\text{closeness}}(v_2) = 1 + 2 + 1 = 4$
\item $C_{\text{closeness}}(v_3) = 1 + 2 + 2 = 5$
\item $C_{\text{closeness}}(v_4) = 1 + 1 + 2 = 4$
\end{itemize}
The centre of the graph is $v_1$.
\end{column}
\begin{column}{0.45\textwidth}
\includegraphics[scale=0.4]{simple_graph.pdf}
\end{column}
\end{columns}

## Closeness Centrality Problem

### Problem

Find the most central vertex, using closeness centrality.

```{r, echo=FALSE, fig.width=1.5, fig.height=1.5}
#g = graph.formula(A-B, A-C, A-D, B-D)
g = erdos.renyi.game(7,p=0.6, directed=FALSE)
par(mar = c(0,0,0,0))
V(g)$label.cex = 0.8
plot(g, vertex.label = c(expression("v"[1]),expression("v"[2]),expression("v"[3]),expression("v"[4]),
expression("v"[5]),expression("v"[6]),expression("v"[7])),
layout=layout.fruchterman.reingold, vertex.size = 35)
```


## Between Centrality


Centrality can be measured by the usefulness of the vertex.  If the
shortest path between vertices $u_i$ and $u_j$ contain vertex $v$,
then the betweenness centrality of $v$ is considered increase.

If we define $\text{sp}_{u_i,u_j}$ as the number of shortest paths
from vertices $u_i$ to $u_j$ and $\text{sp}_{u_i,u_j}(v)$ as the number of
shortest paths from vertices $u_i$ to $u_j$ that pass through $v$,
then the betweenness centrality of $v$ is:
$$
	C_{\text{betweenness}}(v) = {\sum_{u_i, u_j \in V, u_i\ne u_j} \frac{\text{sp}_{u_i,u_j}(v)}{\text{sp}_{u_i,u_j}}}
$$
Note that a shortest path containing $v$ at its start or end cannot
pass though $v$. Therefore, we can ignore all paths that begin with
$v$ and end with $v$.


## Between Centrality Example

### Example

\begin{columns}
\begin{column}{0.6\textwidth}
Shortest paths:
\begin{itemize}
\item $v_1 : v_2$
\item $v_1 : v_3$
\item $v_1 : v_3 : v_4$, $v_1 : v_5 : v_4$
\item $v_1 : v_5$
\item \alert{$v_2 : v_1 : v_3$}
\item \alert{$v_2 : v_1 : v_3 : v_4$, $v_2 : v_1 : v_5, v_4$}
\item \alert{$v_2 : v_1 : v_5$}
\item $v_3 : v_4$
\item \alert{$v_3 : v_1 : v_5$}, $v_3 : v_4 : v_5$
\item $v_4 : v_5$
\end{itemize}
\end{column}
\begin{column}{0.4\textwidth}
%pdf("between_graph.pdf", width=3, height=3)
%g = graph.formula(A-C, A-D, B-D, A-E, B-E)
%par(mar = c(0,0,0,0))
%V(g)$label.cex = 1.4
%plot(g, vertex.label = c(expression("v"[1]),expression("v"[2]),expression("v"[3]),expression("v"[4]),expression("v"[5])), 
%layout=layout.fruchterman.reingold, vertex.size = 30)
%dev.off()
\includegraphics[scale=0.6]{between_graph.pdf}
\end{column}
\end{columns}

$C_{\text{betweenness}}(v_1) = 1/1 + 2/2 + 1/1 + 0/1 + 1/2 + 0/1 = 3.5$ 


## Between Centrality problem

### Problem

Find the most central vertex, using between centrality.

```{r, echo=FALSE, fig.width=1, fig.height=1}
g = graph.formula(A-B, B-C, C-D, D-A, C-E, B-E)
par(mar = c(0,0,0,0))
V(g)$label.cex = 0.8
plot(g, vertex.label = c(expression("v"[1]),expression("v"[2]),expression("v"[3]),expression("v"[4]),expression("v"[5])), layout=layout.fruchterman.reingold, vertex.size = 50)
```




## Eigenvalue Centrality

Eigenvalue centrality computes the centre by stating that the
centrality of a vertex is proportional to the sum of the
centrality of its connected vertices.
$$
	C_{\text{eigen}}(v_i) \propto {\sum_{v_j \in L(v_i)} C_{\text{eigen}}(v_j)}
$$
If we have the adjacency matrix $A$ with elements $a_{i,j}$, the above becomes:
$$
	\lambda C_{\text{eigen}}(v_i) = {\sum_{v_j \in G} a_{i,j} C_{\text{eigen}}(v_j)}
$$
or in vector notation:
$$
	\lambda\vec{c} = A\vec{c}
$$
which is an eigenvalue problem. Note that $\lambda$ is a constant that
allows us to change the $\propto$ to $=$.  The eigenvector $\vec{c}$
of the matrix $A$ contains the centrality scores.  Note that there may
be many eigenvectors of $A$, but only one will contain all positive
elements. We choose this as the solution.


## Use of Centrality

Kevin Bacon is not the centre of the movie universe:

[http://www.buzzfeed.com/tessastuart/40-people-with-more-hollywood-connections-than-kevin-bacon](http://www.buzzfeed.com/tessastuart/40-people-with-more-hollywood-connections-than-kevin-bacon)


# Random Graphs

## Using random graphs

Randomly generated models allow us to examine the behaviour of random
processes. If we obtain a graph from a social network on one day, the
graph might be different the next day. If we can find a random model
that closely follows a random process, we gain insight into the random
process.

## Erdös-Renyi Graph

The simplest random graph is the Erdös-Renyi graph. The parameters of
the random graph are the number of vertices $n$, and the probability of
each edge existing $p$.

- Given $n$ vertices, there are a possible $n(n-1)/2$ edges.
- For each possible edge, we sample a number from the Uniform distribution (between 0 and 1).
- If the random sample is less than $p$, the edge is added.

## Example Erdös-Renyi Graphs ($n = 5$, $p = 0.5$)

```{r, echo=FALSE, fig.width=5, fig.height=2.8}
library("igraph")
par(mar = c(0,0,0,0))
par(mfrow=c(2,3))
for (a in 1:6) {
g = erdos.renyi.game(5,p=0.5)
V(g)$label.cex = 0.8
plot(g, layout=layout.fruchterman.reingold, vertex.size = 35)
}
```

## Barabasi-Albert Graph

Barabasi-Albert random graph is generated in a way that reflects the
growth of many evolving graphs such as the Web and social networks.
The parameters are the number of vertices $n$, the number of edges
to be added per step $e$, and the power $k$.

- Start with one vertex
- Place each vertex one at a time.
- When a vertex is placed it is randomly connected to $e$ existing vertices.
- The probability of connecting to vertex $v_i$ is:

$$
	P(v_i) = \frac{\text{deg}(v_i)^k}{\sum_{v_j \in V} \text{deg}(v_j)^k}
$$
New vertices are more likely to added to vertices with a high degree.

## Example Barabasi-Albert Graphs ($n = 5$, $k = 1$)

```{r, echo=FALSE, fig.width=5, fig.height=2.8}
library("igraph")
par(mar = c(0,0,0,0))
par(mfrow=c(2,3))
for (a in 1:6) {
g = barabasi.game(5,power = 1, m = 1, directed=FALSE)
V(g)$label.cex = 0.8
plot(g, layout=layout.fruchterman.reingold, vertex.size = 35)
}
```

## Example Barabasi-Albert Graph ($n = 100$, $k = 1$)

```{r, echo=FALSE, fig.width=5, fig.height=2.8}
library("igraph")
par(mar = c(0,0,0,0))
par(mfrow=c(1,1))
g = barabasi.game(100,power = 1, m = 1, directed=FALSE)
V(g)$label.cex = 0.5
plot(g, layout=layout.fruchterman.reingold, vertex.size = 10)
```


## Degree Distributions

BA model: Many vertices with low degree, few with high degree.

\vspace{-6em}
```{r, echo=FALSE, fig.width=5, fig.height=4}
plot((1/(1:10))/sum(1/(1:100)), ylim=c(0,0.21), xlab="Degree", ylab="Probability", pch=2, col=2)
points(dbinom(1:10, 99, 0.05), col=1, pch=1)
legend('topright', c("Erdös-Renyi","Barabasi-Albert"), col=c(1,2), pch=c(1,2))

```

## Properties of Social Networks

Social Networks have shown to follow the Barabasi-Albert model, having
a power law distribution (exponentially decreasing) of vertex degree.

Social networks show a few vertices are hubs (have a high degree).

Social networks a robust: if vertices are removed at random, there are
still paths between all vertices. But if carefully selected vertices
(the hubs) are removed, the network will become disjoint.

Social networks form clusters and sub-clusters.

Social networks have small diameters.

The Barabasi-Albert model is a good model for simulating social networks.



## Summary


- Graph theory is essential for analysis of social networks.
- Graphs consist of vertices and edges. They can also be directed and
  weighted.
- There are various subgraphs that we can examine.
- There are various ways to measure centrality.
- We are able to generate random graphs using statistics from real graphs.
- Social networks resemble a Barabasi-Albert random graph.

## Next Week

Visualisation.


