% Introduction to Social Networks
% 300958 Social Web Analytics
% Week 1


# Introduction to the Unit

## Motivation

People are uploading data (text, images, videos, links) to the Web at
every moment of the day. Some of you may be doing it right now!

\vspace{1em}
With proper analysis, we can use this data to examine the state of the
world and predict the future directions of the world's communities.

\vspace{1em}
In this unit, we will learn how to perform this analysis.

## Examining the Learning Guide

The learning guide contains a description of:

- the content of the unit
- what is expected from each student
- delivery of the unit
- the assessment

The learning guide is found in vUWS > 300958 > Unit Information


## Teaching Team

Unit Coordinator:

- Laurence Park <lapark@scem.uws.edu.au>

Lecturers:

- Laurence Park <lapark@scem.uws.edu.au>
- Glenn Stone <g.stone@uws.edu.au>

Tutors:

- Attila Egri-Nagy <a.egri-nagy@uws.edu.au>
- Don Shearman <d.shearman@uws.edu.au>
- Kingsley Agho <k.agho@uws.edu.au>
- Nima Ramezani 
- Shatha Aziz


# Social Networks

## What are Social Networks?

Social Networks consist of a set of items and connections showing how each of
the items interact. We can have a social network of:

- a set of students, where a pair of students is connected if they attend the same class,
- a set of cities, where a pair of cities is connected if share a highway,
- a set of programming languages, where a pair is connected if they share a paradigm,
- a set of people, where a pair is connected if they are related.


Social Network Analysis is the analysis of these networks to obtain
information such as which item is the most influential, and if new items
are introduced how does the network change.

## Small World: Six degrees of separation

The theory that two randomly selected people can find a chain of friends
that is at most six in length.

\vspace{1em}
To examine the small world theory an experiment was devised using the
U.S. Postal Service.

- [https://en.wikipedia.org/wiki/Small_world_experiment]([https://en.wikipedia.org/wiki/Small_world_experiment#The_experiment)


## Paul Erdös


The small world phenomenon extends to many communities. One of the earliest
examined was the academic research community.

- The centre of the research community is thought to be Paul Erdös, who
published about 1500 articles with 509 co-authors.
- Researchers can measure their centrality with respect to this research
social network by computing their [Erdös number](http://www.ams.org/mathscinet/collaborationDistance.html).

### Erdös Number

If you have an Internet connected device, use the above link to examine the Erdös Number of a few of you lecturers.

## Kevin Bacon

Databases such as the [IMDB](http://www.imdb.com/) have allowed us to
examine how connected the set of actors are through the movies they appear
in together.

\vspace{1em}
It was thought that the most connected actor was [Kevin
Bacon](https://en.wikipedia.org/wiki/Kevin_Bacon).

\vspace{1em}
The movie distance between any two actors can be computed at the [Oracle of
Bacon](http://oracleofbacon.org/).

### Bacon Number

If you have an Internet connected device, use the above link to
examine the Bacon Number of a few actors. Can you find any actors with
a Bacon number of 5 or more.

## NSA Surveillance

A recent news article showed that the NSA has put half of the world on
surveillance for suspected terrorist activities ([The
Guardian](http://m.guardian.co.uk/world/2013/jul/17/nsa-surveillance-house-hearing)).



# The Social Web

## The World Wide Web

The World Wide Web is a collection of Web pages and databases that are
served from computers spread across the Internet.

\vspace{1em}
The Web was conceived by Tim Berners-Lee as a method a sharing experimental
particle physics results between labs across the world. It was quickly seen
that this platform could be used to serve all kinds of information.

\vspace{1em}
The Web is intrinsically social in that, for it to grow, it requires us to
upload information that others will read.

## Being Social on the Web

As time has passed, higher level applications have been created for social
activity on the Web:

2. Forums: vUWS Discussion Board, Google Groups
3. Bookmarking: Delicious, Pinterest, Reddit, Slashdot
3. Video: YouTube, Vimeo
4. Images: Flickr, Instagram
7. Music: Last.fm
8. Bloging: LiveJournal, WordPress
9. Academic: Mendeley
9. Business: LinkedIn
5. Location: Foursquare
6. MySpace, Facebook, Google+
5. Microblog: Twitter, Vine, Identi.ca, Tumblr
9. Information seeking: StackExchange


## Twitter

[Twitter](http://twitter.com/) is a Web based social networking
service that allows us to post messages of at most 140 characters, called tweets.

- Twitter is social in that we are able to follow the posts of others
and they are able to follow us.
- The relationships at directional, meaning that $A$ can follow $B$
without $B$ having to follow $A$.
- Everyone's tweets are public and searchable.

Twitter recently released that over 340 million tweets are posted per
day, with over 1.6 billion search queries per day.

## Facebook

[Facebook](http://facebook.com/) is a general social networking
service that allows users to exchange information.

- Facebook is social in that it allows us to be selective of who is
  able to view the information we post.
- The relationships are non-directional, meaning if $A$ is friends
  with $B$ then $B$ is also friends with $A$.
- Specific information in Facebook can be restricted to selected
  friends and therefore not available to the public.

Facebook has over 1 billion users.

## Contributing to the Social Web


For Web based social networks to exist, we must freely contribute and
view information from them. 

\vspace{1em}
The recent introduction of the mobile devices (such as phones and
tablet computers) that have access to the Web has added to the
popularity of Web based social networks. They allow us to conveniently
upload and view information as it happens.

### Our presence on the Social Web

Turn to the people around you and tell them 

- which Social Web services you have contributed information to.
- how often you add and read information from them
- methods which you add and view information from them
- why you add and view information from them

# Analysis of the Social Web

## Information in the Social Web

Many social networking services on the Web have become a place for us to record our lives.

- We post images of events
- We share interesting sites with others
- We write reviews of products
- We broadcast what we have has for breakfast

The social Web has become a place where we feel comfortable releasing
our personal information.

## Using personal information

Businesses have an interest in the Social Web:

- The top priority for all businesses is to make a profit, usually by
offering a product or service.
- Businesses must invest in the development of a product or service
before it can be offered to the public.
- If the public does not buy the product or service, the business had failed.

Analysis of the Social Web allows us to predict the behaviour of the
public under certain situations.

## Getting opinions

Governments have an interest in the Social Web:

- The role of the government is to develop and implement policies and draft laws.
- The polices and laws affect the public.
- The public decide who is voted into government.

Analysis of the Social Web allows us to gauge the opinion of the
public at given times.


# Data Analysis with R

## Introducing R

This semester, we will be using R to perform our data analysis.


## What is `R`?
`R` is a software environment for statistical computing and graphics. It runs on
just about any platform (except iPad!) and is completely free (in the GNU sense).

It is used extensively by academic statisticians for research and teaching and is gaining ground in business.

It has 4634 extension packages available.

### Pros
Its free and open source. It has most methods for most things mostly before any other package. It has the best graphics. It extendable.

### Cons
It has a steep learning curve. No GUI by default. Poor (but improving) memory management; difficulty with very large data sets. 

```{r, echo=FALSE}
opts_chunk$set(prompt=TRUE, comment=NA, dev='tikz')
options(width=54)
```

## `R` Resources
* [http://www.r-project.org](http://www.r-project.org) --- Main `R` website.
* CRAN --- [http://cran.csiro.au](http://cran.csiro.au) --- Comprehensive `R` Archive Network --- base software and add-on packages.
* RStudio --- [http://www.rstudio.com](http://www.rstudio.com) --- is a powerful IDE for `R`
* `R` Commander --- `install.package(Rcmdr)` --- is a partial GUI interface to `R` --- requires TclTk.
* `R` Graph Gallery --- [http://gallery.r-enthusiasts.com/](http://gallery.r-enthusiasts.com/) --- loads of pretty pictures.
* [http://cran.csiro.au/doc/contrib/Torfs+Brauer-Short-R-Intro.pdf](http://cran.csiro.au/doc/contrib/Torfs+Brauer-Short-R-Intro.pdf) --- ``A (very) short Introduction to R''
* ``Introductory Statistics with R'', Peter Dalgaard, Springer 2008.

## `R` Commands
`R` can be used as a basic calculator.
```{r}
1+1
sqrt(2)
2^5
```
It can store things as named objects
```{r}
x <- 1
print(x)
```

## `R` Commands
It understands vectors and matrices
```{r}
x <- c(1,2)
m <- matrix(c(1,2,3,4), ncol=2, byrow=TRUE)
print(m)
m %*% x
```

## `R` Commands
It has functions, and you can write them
```{r}
x <- sqrt(2)
sqr <- function(x) x^2
sqr(2)
```

## Data in R
is stored in `data.frames` 
```{r}
head(iris)

```

## Data in R
```{r}
dim(iris)
```
Some columns are numeric others are factors
```{r}
sapply(iris, class)
```
Data can read from text files (`read.csv` and `read.table`) and various formats
using the `foreign` package.


## Basic Statistics
```{r} 
x <- rnorm(100)
mean(x)
var(x)   ### sd(x)
fivenum(x)
```
minimum, lower-quartile, median, upper-quartile, maximum

## Basic Statistics
```{r} 
summary(iris)
```


## Basic Statistics
```{r} 
t.test(x)
```

## `R` has extensive plotting
```{r, dev=c('pdf'), fig.cap=NULL, fig.width=4, fig.height=2.8, echo=2}
par(pch=16, mar=c(4.1,4.1,1.1,1.1))
plot(Sepal.Length~Sepal.Width,col=Species,data=iris)
legend("topright", col=1:3, legend=levels(iris$Species), pch=16, cex=0.5)
```

## `R` has extensive plotting

```{r, fig.cap=NULL, dev=c('pdf'), fig.width=4, fig.height=2.8, echo=2}
par(pch=16, mar=c(4.1,4.1,1.1,1.1))
plot(Sepal.Length~Species, data=iris)
```



## Summary

- Social Networks describe interaction amongst a set of elements.
- Social Networks appear everywhere.
- The Web has provided us a basis for social interaction.
- Analysing social networks on the Web provides us an insight of the state of the community.


## Next Week

Introduction to R programming and data structures.

