% Introduction to the Social Web
% 300958 Social Web Analysis 
% Week 2 Lab Solutions

```{r setup, include=FALSE}
opts_chunk$set(fig.path='figure/week02-labsol-')
```

`r opts_chunk$set(prompt=TRUE, comment=NA)`
`r #knit_hooks$set(source = function(x, options) paste("<pre lang='r'>", x, "</pre>", sep = ""))`


Copy and past the below code into the RStudio text editor, save it and source it:

```{r eval=FALSE}
main <- function() {
     f = findFactors(1012)
     print(f)
}

findFactors <- function(x) {
   factors = c()
   for (a in 1:x) {
       if ((x %% a) == 0) {
       	  factors = c(factors, a)
       }
   }
   return(factors)
}

main()
```

- What is the code doing?

Two functions are defined (`main` and `findFactors`), then `main` is
called. `main` calls `findFactors` with the argument 1012, then prints the
return value. `findFactors` searches for the set of factors of its
argument `x`, by testing if the numbers from 1 to x divide into x.

- What is the `%%` operator?

This is the remainder operator. E.g. 10 %% 3 gives 1.

- What do you expect the output to be if we change 1012 to 16? Try it to check your answer.

The output will be 1, 2, 4, 8, 16.

- Can you think of a better way to write the `findFactors` function?
  If so, change the function (Hint: factors come in pairs, so we only
  need to check from 1 to $\sqrt{x}$).


```{r}
findFactors <- function(x) {
   factors = c()
   upper = floor(sqrt(x))
   for (a in 1:upper) {
       if ((x %% a) == 0) {
       	  factors = c(factors, a)
       }
   }
   remainingFactors = rev(x/factors)
   if (floor(sqrt(x)) == sqrt(x)) {
       remainingFactors = remainingFactors[-1]
   }
   return(c(factors, remainingFactors))
}

findFactors(16)
```

- Bonus task: Using the function `length`, write a function to test if
  a given number is prime.

```{r}
isPrime <- function(x) {
  if (length(findFactors(x)) == 2) {
    return(TRUE)
  }
  return(FALSE)
}

isPrime(10)
isPrime(19)

```
