% Graphs 1: Graphs and their parameters
% 300958 Social Web Analysis 
% Week 6 Lab

```{r setup, include=FALSE}
opts_chunk$set(fig.path='figure/week06-lab-')
```


# Requirements

For this lab, we will be using the `igraph` package. Install the
package before beginning the lab:

```{r eval=FALSE}
install.packages("igraph")
```

If the package does not download due to the UWS proxy server, see [Lab
5](300958.week05.lab.html) for further information on installing
packages.

Once installed. Load the library:
```{r}
library("igraph")
```


# Creating a graph


To create a graph, we can use the formula interface or provide an
adjacency matrix.

## Graph Formula

We can create a graph by providing the `graph.formula` function with the set of vertices, and how they are connected (edges). For example:
```{r}
g1 = graph.formula(A-B, A-C, A-D, B-D)
```
To view the graph, we can print out the graph variable `g1`:
```{r}
print(g1)
```
To view the vertices of the graph, use the function `V`:
```{r}
V(g1)
```
To view the edges, use the function `E`:
```{r}
E(g1)
```
To visualise the graph, plot it:
```{r fig.cap="The graph `g1`"}
plot(g1)
```

## Adjacency Matrix

To create a graph from an adjacency matrix, we must first create the
matrix. An adjacency matrix $A$ contains $N$ rows and columns, where
$N$ is the number of vertices. The element $a_{i,j}$ of $A$ is the
element from the $i$th row and the $j$th column. If there is an edge
between the $i$th vertex and $j$th vertex, then $a_{i,j} = 1$.
If there is no edge, then $a_{i,j} = 0$.

We can first make a $4\times 4$ matrix, containing all zeros:
```{r}
A = matrix(0,4,4)
```
Then add the edges by allocating ones. We will make the same graph
shown above in `g1`. We want to connect the first vertex to the
second, third and fourth vertices:
```{r}
A[1,c(2,3,4)] = 1
```
We also want to connect the second vertex to the first and fourth:
```{r}
A[2,c(1,4)] = 1
```
We connect the thrid vertex to the first:
```{r}
A[3,1] = 1
```
And connect the fourth vertex to the first and second:
```{r}
A[4,c(1,2)] = 1
```
Giving us:
```{r}
print(A)
```
Notice that the matrix is symmetric. Adjacency matrices for undirected
graphs are always symmetric, showing that the edges can be followed
from either direction.

We create the matrix with:
```{r}
g2 = graph.adjacency(A)
```
and visualise it:
```{r fig.cap="The graph `g2`. The arrow heads on the edges imply that the graph is directed, but it is not."}
plot(g2)
```

Notice that the edges in the plot of `g2` have arrows, implying that
it is a directed graph. Examine the help page of the function 
`graph.adjacency` and work out how to make the graph undirected.

## Edge List

We can also create a graph using an edge list. An edge list is an
$M\times 2$ matrix, containing $M$ edges. Each row of the edge list
provide the start and end vertices. For example:
```{r}
el = matrix(c("A","A","A","B","B","C","D","D"),4,2)
print(el)
```
We then create the graph:
```{r fig.cap="Graph from an edge list"}
g4 = graph.edgelist(el, directed=FALSE)
plot(g4)
```

# Creating Random Graphs

We saw in the lecture that we are able to create an Erdős-Renyi Graph
once given the parameters $n$ (the number of vertices) and $p$ (the
probability of an edge appearing).

To create and Erdős-Renyi Graph:
```{r, fig.cap="Random Erdős-Renyi Graph"}
g.er = erdos.renyi.game(n = 100,p = 0.1)
plot(g.er,  vertex.size = 5)
```

In the plot command, we reduced the vertex size so we can see the
edges more clearly. Run the plot command without `vertex.size = 5` to
see the difference.


To create a Barabási–Albert Graph, we must provide $n$ (the number of
vertices). We can also provide the $k$ (the power) and $m$ (the number
of edges to add to each new vertex).
```{r, fig.cap="Random Barabási–Albert Graph"}
g.ba = barabasi.game(n = 100, directed=FALSE)
plot(g.ba,  vertex.size = 5)
```

# Examining the Graphs

In this section, we will examine some of the functions that that
available for us to examine properties of graphs.

## Density

By visually examining the two graphs above, which looks denser? Use
the function `graph.density` to compute the density of each graph and
compare the results to your guess.


## Diameter

The diameter is the longest shortest path. Which of the two graphs do
you expect to have the largest diameter?  Use the function `diameter`
to compute the diameter of each graph.


## Degree

What do you expect the degree distribution of each graph to look like?
We can compute the degree of each vertex using the function `degree`.
We can also compute the degree distribution of the graph using the function
`degree.distribution`. Use the `help` pages to understand the output.

**Which** vertex is most central according to Degree Centrality?

## Closeness

We defined the closeness centrality of a vertex $v$ as the sum of the
distance from $v$ to all other vertices. To compute the closeness of
each vertex, we use:
```{r}
closeness(g.ba)
```
This does not look right. The sum of differences should be an integer,
but the R output is a set of real values. 

**Read** the R help page for
`closeness` to find what R is computing. Then work out which of the
vertices is the most central with respect to closeness centrality.


## Betweenness

Betweenness centrality measures how often a vertex is used in shortest
paths. We can compute betweenness using:
```{r}
betweenness(g.er)
```

**Is** the centre the same for all three centrality measures? Examine this for the 
Erdős-Renyi graph and Barabási–Albert graph.


# Small Graph

Using the following graph:
```{r fig.cap="Small Graph"}
g3 = graph.formula(A-B, A-C, A-D, B-D, B-E, E-D, C-E)
plot(g3)
```
Calculate the:

- Degree Distribution
- Degree Centrality
- Closeness Centrality
- Betweenness Centrality

using the methods shown in the lecture. Then check your answer using the R functions.


# Twitter Graph

We will now access Twitter to obtain a graph constructed from a *seed*
user.

Load the twitteR library:
```{r}
library("twitteR")
```
and register your application with Twitter using the process from 
[Lab 5](300958.week05.lab.html)

## Dowloading a user

Each twitter user has a set of *friends* and a set of *followers*:

- Followers of user $x$: users who are following $x$.
- Friends of user $x$: users who $x$ is following. 

Therefore, a user can choose their friends, but cannot choose their
followers. There are many users of Twitter, we want to find the
interesting ones. Interesting users usually have many followers
(because they are interesting). So when obtaining information about
users on Twitter, we should note:

- Popular people have many followers.

Using `twitteR`, we can download user information when given a screen
name or ID.
Let's examine [Wil Wheaton](http://en.wikipedia.org/wiki/Wil_wheaton):
```{r eval=FALSE}
user = getUser("wilw")
```
The variable `user` is now a `twitteR` object, which we can use as a seed for our graph.

To examine all of the details for Wil Wheaton, convert the twitteR object into a data frame:
```{r eval=FALSE}
user$toDataFrame()
```
The above function provides information such as the number of friends,
the number of followers, if the account is protected and verified and
the owner's name and id.  Wil Wheaton is a popular person, so he must
have popular friends. From the output of `user$toDataFrame()` identify
how many friends Wil Wheaton has.

## Dowloading a user's friends.

We now want to download the list of Wil Wheaton's friends. Since he has
a large number of friends, we download the first 100.
```{r eval=FALSE}
friends = user$getFriends(100)
```

The above function gets the friends of `user`; since `user` contains
Wil Wheaton's details, the function returns a list of his friends.  The
list of friends is a list of `twitteR` objects, so `friends[[1]]` is the
same data type as `user`.


We now have the details of 100 of Wil Wheaton's friends. 
Examine the information of the first friend:
```{r eval=FALSE}
friends[[1]]$toDataFrame()
```
We can find which friends are the most popular by examining their follower count.
Here we examine the follower count of the first friend:
```{r eval=FALSE}
friends[[1]]$getFollowersCount()
```
To examine the follower count of all friends:
```{r eval=FALSE}
follower.count = c()
for (a in c(1:length(friends))) {
    follower.count[a] = friends[[a]]$getFollowersCount()
}
```
Let's make this a function for convenience:
```{r eval=FALSE}
count.followers = function(friends) {
    follower.count = c()
    for (a in c(1:length(friends))) {
        follower.count[a] = friends[[a]]$getFollowersCount()
    }
    return(follower.count)
}
```

**Find** the 10 friends that have the most followers. What are their
names?  Note the function `sort` will sort the vector of follower
counts. The function `order` will sort, but provide the position of
the sort. So to find the top 10, we use `order` with `decreasing=TRUE`
and choose the first ten values, giving us the positions of the top
10.


Some Twitter accounts are protected, meaning that we can't read them
unless we are a follower. We can detect if the first friend account is protected using:
```{r eval=FALSE}
friends[[1]]$getProtected()
```

**Write** a `for` loop to check if any of the friend accounts are
protected and store the `TRUE`/`FALSE` values in the variable `protected.status`. 

Once we have the vector `protected.status`, we can extract the unprotected friends of Wil Wheaton:
```{r eval=FALSE}
unprotected.friends = friends[!protected.status]
```
This will select the positions that are FALSE (the exclamation mark !
changes FALSE to TRUE and TRUE to FALSE).

**Using** the function `count.followers` find the 10 unprotected
  friends with the most followers and store them in `top.unprotected.friends`.


Now that we have the list of top 10 unprotected friends, we can use this to download the
friends' friends.
Lets create a place to store the friends of friends.
```{r eval=FALSE}
more.friends = list()
```
We can download a list of 100 friends of the first friend of Wil Wheaton using:
```{r eval=FALSE}
more.friends[[1]] = top.unprotected.friends[[1]]$getFriends(100)
```

**Write** a `for` loop to download 100 friends from the 10 most
popular friends of Wil Wheaton and store them in `more.friends`.


## Creating the Twitter Graph

Now we have Wil Wheaton, 100 of Wil Wheaton's friends, and 100 of 10 of
Wil Wheaton's most popular friends.

We can create a graph by constructing an edge list (who links to who).
We know Wil Wheaton links to all of his 100 friends, so the edge list
will contain 100 rows beginning with Wil Wheaton and ending with the
friend of Wil Wheaton. First we must get the names of all of Wil Wheaton's friends. 

First we allocate a variable to contain the names:
```{r eval=FALSE}
friend.names = c()
```
We can get the first name using:
```{r eval=FALSE}
friend.names[1] = friends[[1]]$getScreenName()
```

**Write** a `for` loop to store all 100 screen names in the variable `friend.names`.

We can now build the edge list using:
```{r eval=FALSE}
wil = rep(user$getScreenName(), length(friends)) # repeat Wil Wheaton's user name 100 times
el = cbind(wil,friend.names) # bind the columns to create a matrix
```

We need to do the same for Wil Wheaton's 10 most popular friends. To
simplify the process, we should write a function. 

**Using** what you have done above, write the function:
```{r eval=FALSE}
user.to.edgelist <- function(user, friends) {

  # create the list of friend screen names
  friend.names = c()		
  for (a in c(1:length(friends))) {
      # enter missing code
  } 

  user.name = rep(user$getScreenName(), length(friends)) # repeat user's name
  el = cbind(user.name,friend.names) # bind the columns to create a matrix

  return(el)
}
```

We can use the created function `user.to.edgelist` to create the edge list for Wil Wheaton:
```{r eval=FALSE}
el.wil = user.to.edgelist(user, friends)
```
We can also build the edge list for the top 10 friends using a loop:
```{r eval=FALSE}
for (a in c(1:length(more.friends))) {
    el.friend = user.to.edgelist(top.unprotected.friends[[a]], more.friends[[a]])
    el.wil = rbind(el.wil, el.friend) # append the new edge list to the old one.
}
```
Now that we have the edge list, we can create the graph:
```{r eval=FALSE}
g = graph.edgelist(el.wil)
```
Let's plot the graph. Since there are many vertices, we will reduce
the vertex size and use a spacial plot layout:
```{r eval=FALSE}
plot(g, layout=layout.fruchterman.reingold, vertex.size = 5)
```
This graph contains many vertices that we did not examine. To remove
these, let's only keep the vertices with degree (in or out) greater
than 1.
```{r eval=FALSE}
g2 = subgraph(g, which(degree(g, mode="all") > 1))
```
This graph is now easier to visualise:
```{r eval=FALSE}
plot(g2, layout=layout.fruchterman.reingold, vertex.size = 5)
```

**Who** is at the centre of the graph? Use the centrality measures to
  examine this.

**Examine** the graph density. Is it sparse or dense?

**Examine** the degree distribution. Is this graph more similar to an
Erdős-Renyi graph or a Barabási–Albert graph?
