% Simple Exposure Analysis
% 300958 Social Web Analysis 
% Week 4 Lab Solutions


```{r setup, include=FALSE}
opts_chunk$set(fig.path='figure/week04-labsol-')
```

`r opts_chunk$set(prompt=TRUE, comment=NA)`

# Using R

- Use `read.csv` to read in the CSV files [keyMetrics.csv](keyMetrics.csv), [LifetimeLikesByGenderAge.csv](LifetimeLikesByGenderAge.csv) and [WeeklyReachDemog.csv](WeeklyReachDemog.csv) into data frames.
Make sure you give sensible names to the data frames you create. Use as.is=TRUE for keyMetrics.csv to avoid the mentioned problem.

```{r}
keyMetrics = read.csv("http://staff.scm.uws.edu.au/~lapark/300958/labs/keyMetrics.csv", as.is=TRUE)
LifetimeLikesByGenderAge = read.csv("http://staff.scm.uws.edu.au/~lapark/300958/labs/LifetimeLikesByGenderAge.csv", as.is=TRUE)
WeeklyReachDemog = read.csv("http://staff.scm.uws.edu.au/~lapark/300958/labs/WeeklyReachDemog.csv", as.is=TRUE)
```

- Identify and extract the "Weekly Total impressions" variable from the keyMetrics. HINT: the `names` function will
give a list of column names.

```{r}
names(keyMetrics)
impress <- as.numeric(keyMetrics[-1,28])
print(impress)
```


- Look at the help page for `plot` (see the examples) 
and try changing the axis labels (`xlab=` and `ylab=`), give a title (`main=`) and
changing colours and line type (`col=` and `lty=`)

```{r}
dates <- strptime(keyMetrics[-1,1], format="%m/%d/%y")
plot(dates,impress, type="l", col=2, lty=3)
```


# $\chi^2$ Test


## Test for independence

Create the two column table.
```{r}
WeekReach = read.csv("http://staff.scm.uws.edu.au/~lapark/300958/labs/WeeklyReachDemog.csv", as.is=TRUE)
tab = matrix(as.numeric(WeekReach[158,3:16]), nrow=2, byrow=TRUE)
colnames(tab) <- c("13-17", "18-24", "25-34", "35-44", 
                    "45-54", "55-64", "65+")
rownames(tab) <- c("Female","Male")

stretchTable = function(tab, variableNames) {
    tabx = rep(rownames(tab), rowSums(tab))
    l = ncol(tab)
    m = nrow(tab)
    cn = colnames(tab)
    taby = c()
    for (a in 1:m) {
        for (b in 1:l) {
            taby = c(taby, rep(cn[b], tab[a,b]))
        }
    }

    d = data.frame(x = tabx, y = taby)
    colnames(d) = variableNames
    return(d)
}

tab2 = stretchTable(tab, c("Gender","Age"))
    
# Verify that we the correct values
table(tab2)
```


Compute the $\chi^2$ randomisation distribution.
```{r}
## define the functions to use
expectedIndependent = function(X) {
    n = sum(X)
    p = rowSums(X)/sum(X)
    q = colSums(X)/sum(X)
    return(p %o% q * n) # outer product creates table
}

chiSquaredStatistic = function(X, E) {
    return(sum((X - E)^2/E))
}

## compute the expected table
E = expectedIndependent(table(tab2)) # compute expected counts if independent

## compute the randomisation distribution
x2dist = replicate(1000, { # compute 1000 randomised chi-squared statistics
  ageShuffle = sample(tab2$Age)
  genderShuffle = sample(tab2$Gender)
  Xindep = table(genderShuffle, ageShuffle)
  chiSquaredStatistic(Xindep, E)
})

hist(x2dist)
```

Compute the $\chi^2$ statistic for `tab`.
```{r}
x2 = chiSquaredStatistic(table(tab2), E)
```

Compute the $p$ value of the test.
```{r}
## pval is the proportion of x2dist that is greater than x2
pval = mean(x2dist > x2)
print(pval)
```

The $p$-value is large, so we cannot reject $H_0$. Our conclusion is
that we don't have enough evidence to show that age and gender are
dependent.

## Test for preference

```{r}
age = colSums(tab)
n = sum(age)
k = length(age)
```

Given the sample size `n` and the age table, compute the expected
age group frequencies if all ages have the same proportion.
```{r}
## If all have the same proportion and the proportions must sum to 1:
ep = rep(1/k, k)
## We want the expected frequencies
E = ep*n
```

Compute the $\chi^2$ value using the observed and expected frequencies.
```{r}
x2 = chiSquaredStatistic(age, E)
print(x2)
```

- Sample from the set of age categories with replacement using the given
  proportions and sample size (using `sample`).
- Compute the $\chi^2$ value of the sample.
- Repeat at least 1000 times to obtain a distribution of $\chi^2$ values given $H_0$.

```{r}
x2dist = replicate(1000, {
	r = table(sample(k, size = n, prob = ep, replace = TRUE))
	chiSquaredStatistic(r, E)
})
hist(x2dist)
```

Compute the $p$ value.
```{r}
pval = mean(x2dist > x2)
print(pval)
```
The $p$-value is very small, so we reject $H_0$, meaning that we have evidence that the ages are not of equal proportion.


# More R - a challenge

- Construct the $2\times 2$ table containing the number of males and females
(independent of age), versus the months of April and May. Does a $\chi^2$ test
show that the reach for gender is independent of the two months?


```{r}
dates = strptime(WeeklyReachDemog$Date, format="%m/%d/%y")
months = format(dates, "%b")
female = rowSums(WeeklyReachDemog[,3:9], na.rm=TRUE)
male = rowSums(WeeklyReachDemog[,10:16], na.rm=TRUE)
X = matrix(c(sum(female[months == "Apr"]),
      sum(female[months == "May"]),
      sum(male[months == "Apr"]),
      sum(male[months == "May"])), 2, 2)
colnames(X) = c("Female", "Male")
rownames(X) = c("April", "May")
print(X)
```

Create the two column single entry table
```{r}
X2 = stretchTable(X, c("Months", "Gender"))
    
# Verify that we the correct values
table(X2)
```

```{r}
## compute the expected table
E = expectedIndependent(table(X2)) # compute expected counts if independent

## compute the randomisation distribution
x2dist = replicate(1000, { # compute 1000 randomised chi-squared statistics
  genderShuffle = sample(X2$Gender)
  monthShuffle = sample(X2$Month)
  Xindep = table(monthShuffle, genderShuffle)
  chiSquaredStatistic(Xindep, E)
})

hist(x2dist)
```

Compute the $\chi^2$ statistic of our original sample.
```{r}
Xstat = chiSquaredStatistic(X, E)
```

Compute $p$-value.
```{r}
pval = mean(x2dist > Xstat)
print(pval)
```
The $p$-value is large, so we don't reject $H_0$. We don't have enough evidence to say that month and gender are not independent.


Use the R $\chi^2$ test and compare the $p$-values to ours.
```{r}
chisq.test(X)
```
