% Looking for Trends 
% 300958 Social Web Analysis 
% Week 11 Lab Solutions

```{r setup, include=FALSE}
opts_chunk$set(fig.path='figure/week11-labsol-')
```

`r opts_chunk$set(prompt=TRUE, comment=NA)`

- Complete the following R-code to write a function to compute Twitter's $\chi^2$ statistic for trends.

```{r eval=FALSE, prompt=FALSE}
TwitterChi = function(tweetsBefore, tweetsNow, N) {

  return((tweetsNow - tweetsBefore)^2/tweetsBefore
    + ((N - tweetsNow) - (N - tweetsBefore))^2/(N - tweetsBefore))
  
}
```



- Using the data below (from lectures) compute $\bar{x}$, $\bar{y}$, $SS_X$, $SS_{XY}$ and therefore $\hat{\beta}$ and $\hat{\alpha}$. You should use, $x=\text{day}$ and  $y=\sqrt{\text{count}}$.

```{r}
count = c(135, 145, 133, 102, 105, 108, 128, 144, 
149, 130, 107, 106, 83, 117, 123, 104, 116, 127, 75, 128, 136, 
145, 120, 127, 110, 109, 122, 136, 125, 143, 177, 142, 134, 153, 
173, 151, 174, 168, 158, 156, 128, 156, 140, 133, 132)
day = c(210, 
211, 212, 213, 214, 215, 216, 217, 218, 219, 220, 221, 222, 223, 
224, 225, 226, 227, 228, 229, 230, 231, 232, 233, 234, 235, 236, 
237, 238, 239, 240, 241, 242, 243, 244, 245, 246, 247, 248, 249, 
250, 251, 252, 253, 254)

x = day
y = sqrt(count)
xbar = mean(x)
ybar = mean(y)
SSx = sum((x - xbar)^2)
SSxy = sum((x - xbar)*(y - ybar))
b = SSxy/SSx
a = ybar - b*xbar
print(c(a,b))
```

- Use the `lm` function to check your results on the Twitter counts from above.

```{r}
summary(lm(y ~ x))
```

The following is a moving average (as defined in lectures) function using a `for` loop.

```{r prompt=FALSE}
moving.average = function(x, m) {
  if( m%%2 ) { 
     ## m is odd
     w = rep(1,m)/m
   } else { 
    ## m is even
    w = c(0.5, rep(1,m-1), 0.5)/m
  }
  j = floor(m/2)
  offsets = (-j):j
  n = length(x)
  res = rep(NA, n)
  for(i in (j+1):(n-j)) {
    res[i] = sum(w*x[i+offsets])
    }
  return(res)
}
```

- Make sure you understand this function and use it to compute the 7 point moving average of the tweet count data. Remember to square root first.

```{r}
trend = moving.average(sqrt(count), 7)
print(trend)
```

The following function computes a seasonal component by averaging over all observations that are $m$ apart.

```{r prompt=FALSE}
seasonal = function(x ,m) {
  ## extend the data so it is a multiple of m long
  tmp = c(x, rep(NA, m - length(x)%%m))
  ## convert to a matrix
  mat = matrix(tmp, nrow=m)
  ## Calculate the row means to get the seasonal component (excluding missing entries)
  seas = rowMeans(mat, na.rm=TRUE)
  seas = seas - mean(seas)
  return(seas)
}
```

- Make sure you understand this function and use it to compute the seasonal component for the tweet count data. Remember to subtract the trend first.

```{r fig.cap="Data split into trend and seasonal components."}
centred = sqrt(count) - trend # remove trend
zcentred = centred[!is.na(centred)] # remove NAs at boundaries
s = seasonal(zcentred, 7) # compute seasonal component
print(s)
plot(day, sqrt(count))
lines(day, trend, col=2)
lines(day, trend, col=2)
lines(day[!is.na(centred)], trend[!is.na(centred)] + rep_len(s, length.out = sum(!is.na(centred))))
```


 
