% Introduction to the Social Web
% 300958 Social Web Analysis 
% Week 2 Lab

```{r setup, include=FALSE}
opts_chunk$set(fig.path='figure/week02-lab-')
```

`r opts_chunk$set(prompt=TRUE, comment=NA)`
`r #knit_hooks$set(source = function(x, options) paste("<pre lang='r'>", x, "</pre>", sep = ""))`

# The Social Web

Throughout this semester, we will be making use of Twitter to analyse
social interaction.  To begin this lab, we will create a Twitter
account, add content to the account and follow others in the class and
in the World.


## Creating a Twitter account

To create a Twitter account, go to this URL:
<https://twitter.com/> and fill in your details in the "Sign Up" box.
Make sure to read the Terms and Conditions carefully.

## Twitter Terminology

Twitter is a micro-blogging service that allows its users to write
messages of up to 140 characters in length.  Its usefulness comes from
its ability to allow us to follow others we are interested in, making
it act as a customised news service.  Many posts contain a summary
with a link to a more complete story, allowing us to quickly flick
through the headlines and read any posts that interest
us, in more detail.

The components of Twitter are:

- tweet: a message post
- handle: a Twitter id, beginning with @
- hashtag: any term beginning with #. Hashtags have become Twitter index terms, allowing us to easily find tweets on certain topics.
- follower: an account that is following your posts
- friend: an account that you are following
- feed: the set of tweets
- retweet: adding a previously published tweet to your own feed, allowing your followers to see it
- reply: create a tweet containing another users handle, making it appear in both feeds.


## Following your classmates

Before we can follow others in the class, we have to be able to find
them in Twitter. The simplest way of finding each other is to tweet
something that other can find and then follow you.

Once you have an account, write a tweet containing the tag:

```
#UWS300958
```

For example, I could write:

```
Searching for the existence of life on planet #UWS300958
```

We can then use the twitter search box to search for `#UWS300958` and
find others in the class.

We will be using the data from your twitter accounts later in the unit.
So try to get as many followers and friends as you can.


## Writing a second tweet.

No twitter account could be complete without a tweet linking to a cat
video.  For your second tweet, find a cat video on youtube and write a
tweet containing the cat video URL.

After uploading the tweet, examine the URL in the tweet. Is it the
same as the URL you provided?


## Receiving interesting tweets

The final part of setting up your twitter account is to add sources
that will provide you with information that you are interested in.

The two most common ways of finding twitter accounts to follow are:

- Search for topics that you are interested in (this usually provides
  lots of rubbish, but you may find what you want).
- Examine what your friends and followers are following.




# Setting up R and learning the basics

For the second part of the lab, we will set up R, the programming tool
we will be using through the course.


## Installing R

All of the lab sessions for 300958 Social Web Analysis will use R and
RStudio. These two programs should be installed on the lab
machines. If they are installed, you should be able to click on the
Start menu, type RStudio and the program should appear. If these two
applications are *not* installed on the computer you are using,
download them from:

- R: [http://cran.csiro.au/](http://cran.csiro.au/)
- RStudio: [http://www.rstudio.com/ide/download/desktop](http://www.rstudio.com/ide/download/desktop)

Install R, then Rstudio. Note that R is the program that runs the R
programs. RStudio is an IDE for R that provides a pretty interface for
us to use.


Open RStudio and say hello.
```{r}
print("Hello world!")
```

## Getting familiar with RStudio

Once RStudio is loaded, we will see four sections:

- Text editor (top left). Use this to enter code and save it as a script.
- Console (bottom left). Code can be entered straight into the console, or scripts can be run by entering source("script_filename.r")
- Workspace (top right). This section shows the variables that are currently defined. 
- Browser (bottom right). We can access files, view plots, examine packages and access help in the browser.

We can write code into the text editor or the console. The choice of
where you write code depends on what the code is for.

- The console runs the code we enter, but does no keep it. This is
  useful for testing a line of code, installing packages or looking up
  help files.

- The text editor does not run the code we type, but allows us to save
  the code for later use. There are buttons above the text editor that
  allow us to run the code a line at a time, or run the whole
  script. Most of the lab work we do should be done in the text
  editor, so we can save it, run it, and review it later.


To write your first script, type:
```{r eval=FALSE}
print("Hello world!")
```
into the text editor, save it as hello.r. Then go to the console section and type:
```{r eval=FALSE}
source("hello.r")
```
Rather than typing the above command into the console, there is also a
button that we can use, above the text editor, that sources the text file.

By sourcing the script, you will run every command. Since there is one
command in the script, only one command will be run.

If R complains that it cannot find the file, it may be looking in the wrong directory. 
To examine which directory R is sourcing from, use:
```{r eval=FALSE}
getwd()
```
To set the directory to the directory containing the script, use:
```{r eval=FALSE}
setwd("directory")
```
where `directory` is where your script is stored.


## Errors in code

If there is an error in your program, the error message will be shown
in the console. The error messages are usually descriptive enough for
us to work out where the error is located in out code. **Read error
messages carefully!** The more you get used to the types of errors,
the easier it will be to debug you code.

To see a simple error message, type in the console:
```{r}
hello
```
R will complain that `hello` can't be found. This is because R does
not know what `hello` is. If we first create the variable called `hello`:
```{r}
hello = 10
```
R now knows of an object called `hello` and so typing `hello` in the
console will no longer result in an error. We often see the
previous error message when we have misspelled a variable name.
For example, `hello` is not the same as `Hello`.

If you see an error message, or set of error messages, start from the
first message, fix the bug and run the code again. If you can't
understand what the error is, ask the tutor.


## Installing the required R packages

This semester we will be using a set of R packages that are not
included in the base install.

Due to the current design of the UWS computer labs, if we install any
packages in the default directories, the packages will be deleted
overnight. Therefore, we should tell R to install the packages within
our home directory.

Your home directory is:

- My Documents (for Windows users)
- /Users/username (for OS X, where username is your username)
- /home/username (for Linux, where username is your username)

To tell R where to install the packages

1. In your home directory, create the directories `R/library`, to hold the packages.
2. In your home directory, create the file `.Renviron' in a text editor and add the line:

```
R_LIBS_USER="~/R/library"
```

3. Save the file and restart RStudio

From now on, when the command `install.packages` is issued, the
packages will be stored in the path you assigned.

Try this out by installing the MASS package:
```{r eval=FALSE}
install.packages("MASS")
```
Once the package is installed, check the directory that you created to
make sure that the package was installed in that directory.


## Setting the Proxy Information

If the packages failed to install, read the error message that was
reported in the console.  If the message states that the repository
could not be contacted, we may need to tell R about the UWS network
proxy server.

Information on how to tell R about the proxy is here:

- <https://support.rstudio.com/hc/en-us/articles/200488488-Configuring-R-to-Use-an-HTTP-Proxy>


## Manual download

If the installer fails to fetch the packages, even after providing the
proxy server information, we can download the packages to a local
drive and install them from there.

The list of all packages is here:

- <http://cran.r-project.org/web/packages/available_packages_by_name.html>

From the package list, click on the wanted package to go to the
package page.  Then download either the MacOS or Windows binary to a
local directory.  To install the package, we must tell R where it is
saved. For example, if we downloaded the `tm` package and saved it in
the /tmp directory:
```{r eval=FALSE}
install.packages("/tmp/tm_0.5-9.1.zip", repos=NULL)
```
The argument `repos=NULL` tells the installer not to download from the
R repository.

## Packages still won't install

If you have tried each of the methods above and consulted with the
tutor, but still cannot install the packages, there may be a problem
with the UWS network or a change in the packages. To let Laurence know
about this, create an issue at the 300958 issue tracker.

- <https://bitbucket.org/lapark/300958-social-web-analytics/issues>

If an issue already exists containing the same problem, then someone
else has reported it and there is no need to create a new issue.


## Loading packages

Installing packages only downloads the appropriate package for your machine.
To use the package, we must load it (just as we use `include` in C and C++
and `import` in Python).

To load a package, we use the command `library`. To load the package MASS, we use:

```{r eval=FALSE}
library("MASS")
```

Note that we cannot load packages that we have not already
installed. Loading a package reads the package files and stores them
in the computers memory, to make sure that the functions and data they
offer are available to use.

The packages that we will be using this semester are:

- igraph
- maps
- ROAuth
- skmeans
- spatstat
- tm
- twitteR
- wordcloud

Install these packages now, so that they are ready to be used in the later labs.

## Package Contents

To read the package description and examine the functions and data it
offers, we first load the package (using the `library` command), then
we use the R help command.

To read about the package MASS, use:

```{r eval=FALSE}
help(package = "MASS")
```

Read the description of the packages we installed in the previous
section.


## A sample script

Copy and past the below code into the RStudio text editor, save it and source it:

```{r eval=FALSE}
main <- function() {
     f = findFactors(1012)
     print(f)
}

findFactors <- function(x) {
   factors = c()
   for (a in 1:x) {
       if ((x %% a) == 0) {
       	  factors = c(factors, a)
       }
   }
   return(factors)
}

main()
```

- What is the code doing?
- What is the `%%` operator?
- What do you expect the output to be if we change 1012 to 16? Try it to check your answer.
- Can you think of a better way to write the `findFactors` function?
  If so, change the function (Hint: factors come in pairs, so we only
  need to check from 1 to $sqrt{x}$).
- Bonus task: Using the function `length`, write a function to test if
  a given number is prime.

## Final Task

Before the end of the lab, post a tweet on how the lab went.
