% Introduction to R
% 300958 Social Web Analysis 
% Week 3 Lab

```{r setup, include=FALSE}
opts_chunk$set(fig.path='figure/week03-lab-')
```

`r opts_chunk$set(prompt=TRUE, comment=NA)`
`r #knit_hooks$set(source = function(x, options) paste("<pre lang='r'>", x, "</pre>", sep = ""))`


# R Documentation

If you have never programmed in R before, download this document as a
reference:

- [http://cran.csiro.au/doc/contrib/Torfs+Brauer-Short-R-Intro.pdf](http://cran.csiro.au/doc/contrib/Torfs+Brauer-Short-R-Intro.pdf)

Read through it and try some of the examples.

## The R help system

R also has an extensive help system. Each function is documented and
accessible using the function `help`. For example, the documentation
of the function `plot` can be read by issuing the command:

```{r eval=FALSE}
help(plot)
```
or
```{r eval=FALSE, tidy=FALSE}
?plot
```

Each help page contains the sections:

- Description: explaining what the function does.
- Usage: showing how to call the function and what the default arguments are.
- Arguments: Describing the arguments in detail.
- Details: explains how the function behaves.
- See Also: other similar functions.
- Examples: examples of using the function.

The examples section at the bottom of the help page can be very useful
if you have not used the function before.

## Exercise

Go to the help page for `plot` and try out the examples from the
Examples section of the help file.


# Coding in R

The real power of R comes from its ability to run complex programs
with a few lines of code. In this section, we will examine the
fundamentals of coding in R.

## Variables

In R variables are defined once data is assigned. R has dynamic
typing, meaning that we *do not* have to declare the type of the
variable (e.g. we do not have to declare `int x;` as done in C, C++
and Java), the type of the variable is set once data is assigned. To
create a variable `x` of type numeric containing the value `4`, we
issue the command:
```{r}
x = 4
```
If `x` did not exist, R creates the variable and then assigns the value.

We can create variables of type `string`:
```{r}
x = "hello, this is some text."
```
We can also determine the type of the variable using the `class`:
```{r}
class(x)
```


## Data Structures

R provides a few fundamental data structures that can be used by
themselves, or used as building blocks for more complicated structures.


### Arrays

Arrays are set of values that can be accessed by providing the correct
index number.


A one dimensional array is a vector. To create a vector, we use the function `c`:
```{r}
ages = c(24, 43, 76, 21, 32)
```
A two dimensional array is created using the function `matrix`:
```{r}
cross.ages = matrix(c(24, 43, 76, 21, 32, 54), ncol = 2)
```

Higher dimensional arrays can be constructed using the function
`array`. Here is an example of a three dimensional array, where each
dimension has two items:
```{r}
really.cross.ages = array(c(24, 43, 76, 21, 32, 54, 43, 65), c(2,2,2))
```
To examine the variables types:
We can also determine the type of the variable using the `class`:
```{r}
class(ages)
class(cross.ages)
class(really.cross.ages)
```

To access data from an array, we provide the index on the data.
To access the second item of the vector `ages`:
```{r}
ages[2]
```
To access the second and fifth items of the vector `ages`:
```{r}
ages[c(2,5)]
```
To access the second up to the fifth items of the vector `ages`:
```{r}
ages[2:5]
```
To access the first row and second column of the matrix `cross.ages`:
```{r}
cross.ages[1,2]
```
To access the second column of the matrix `cross.ages`:
```{r}
cross.ages[,2]
```

Note that data types cannot be mixed in an array (e.g. it cannot
contain both numeric and string value, only one or the other).


### Data Frames

A data frame is a rectangular table of data. with named columns.  The
`iris` data frame is part of the core R libraries and so is always
accessible in R. Details is the `iris` dataset 
<https://en.wikipedia.org/wiki/Iris_flower_data_set>

We can determine if a variable is a data frame:
```{r}
class(iris)
```
We can examine the size of the data frame:
```{r}
dim(iris)
```
We can examine the names of the data frame columns:
```{r}
names(iris)
```
We can select a column:
```{r}
iris$Petal.Width
```
We can view the first few items of the data frame:
```{r}
head(iris)
```
We can also access the elements using indexing (as if it were a matrix):
```{r}
iris[3,]
```

If we want to find the rows that have a spefic quality, we must use a
truth statement. To find all rows that correspond to the "setosa"
species:
```{r}
iris$Species == "setosa"
```
The set of `TRUE` and `FALSE` values corresond to each row. To find which are `TRUE`, we use the function `which`:
```{r}
which(iris$Species == "setosa")
```
We can store these positions and use them to extract the petal width for all of the Setosa species.
```{r}
pos = which(iris$Species == "setosa")
iris$Petal.Width[pos]
```


To create a new data frame, we provide the column data, where each
column of a vector of the same length:
```{r}
my.data = data.frame(colour = c("red", "green", "blue"), length = c(1,4,3))
```

### Lists

Lists are generic storage data structures that are constructed in a
similar fashion to a data frame, but there is no restriction on the
data they can hold.

We can have a list that stores a persons information:
```{r}
bob = list(name = "Bob Bobson", age = 42, runs = c(13, 45, 87, 2, 9))
```
and examine its type:
```{r}
class(bob)
```
We can even have a list of lists:
```{r}
bobs = list(list(name = "Bob Bobson", age = 42, runs = c(13, 45, 87, 2, 9)),
     list(name = "Bob Johnson", town = "Bobville", bowling.runs = c(45, 23, 54, 12), 
     	       bowling.wickets = c(2,0,2,3)))
```
which still has the same type:
```{r}
class(bobs)
```

There are two methods of accessing a list, single and double
brackets. The single brackets method returns a list contain the items:
```{r}
bobs[1]
```
The double brackets method returns the contents:
```{r}
bobs[[1]]
```
The method chosen is based on what will be done with the data in the
next stage of your program.

If the list contains names, we can access the data by calling the name
(just as in the data frame section):
```{r}
bob$runs
```



### Exercise

What data structure would you choose to store your past results at
UWS? Write code to store your results in the variable `past.results`.


## Flow Control

Code in R can be written in a procedural style or object oriented
style. If a script is run, it is interpreted from the top of the file to the bottom.
The control of the flow can be obtained by writing conditional statements (if, else)
and loops (for, while).

### if statement

If the execution of code depends on the state of a variable, we can
use an if statement. For example, if the operating system is Windows,
the file path is "C:\\User\\greg\\file.txt", else the file path is
"/home/greg/file.txt".

The if statement in R has the syntax:
```{r eval=FALSE}
if (condition) {
   ## run this code if condition is true
}
```
The statement can also be extended to include other conditions and a default condition.
```{r eval=FALSE}
if (condition_1) {
   ## run this code if condition 1 is true
} else if (condition_2) {
   ## run this code if condition 1 is false and condition 2 is true
} else {
   ## run this code if condition 1  and condition 2 are false
}
```


The `if` statement depends on a condition, where a condition is simply
a statement that returns a `TRUE` or `FALSE` value. 
If the variable x is defined as:
```{r}
x = 4
```
A condition could be as simple as:
```{r}
x == 2
```
or it could be a function that exists or we have written:
```{r eval=FALSE}
do.i.like.this.number(2)
```


### for loop

Loops allow us to execute a piece of code a number of times, where the
number is not known until run time.

```{r eval=FALSE}
for (variable in sequence) {
  ### run this code using variable
}
```


For example, for every year that Greg has been born, give him a
birthday punch (remember that `1:5` is the vector `c(1,2,3,4,5)`).
```{r eval=FALSE}
for (year in 1:gregsAge) {
  punchGreg()
}
```

We can also use for loops to run a function over a set of items. For
example, if we have the set of words `words`, where `words[i]` is the
$i$th word, we can count the length of each word using:
```{r eval=FALSE}
for (i in 1:length(words)) {
  countLetters(word[i])
}
```
Note that in this case, we have to write the function `countLetters`.


### Exercise

The function `cat` will print its contents to the screen. E.g.
```{r}
a = 5
cat("This is a number ", a, "\n")
```
Write a for loop to convert a number to a sequence of plus signs. For
example, 5 will be converted to "+++++".


Write an `if` statement to test if an item in `past.results` is a pass
or fail grade.  Then use a for loop to test all of the grades in
`past.results`.



## Writing functions

If a particular piece of code is run often, it may be wise to place it
in a function. In R, functions are stored in variables. If we define the function:
```{r}
addOne = function(x) {
       return(x + 1)
}
```
we can examine the function by typing its name in the console and pressing enter.
To run the function, we must add parentheses to the end and include any arguments:
```{r}
addOne(5)
```
```{r}
x = addOne(5)
addOne(x)
```
If the result is not assigned to a variable, it is printed to the screen.


### Exercise

Write a function `check.pass` that takes the data structure
`past.results` and returns a vector showing if each grade was a pass
or a fail.

# Challenge 1: Finding Force of Will

The file [AllSets.json]("http://x") contains details of all "Magic: The
Gathering" cards, but it is encoded as a JSON file.

The library `jsonlite` contains the function `fromJSON` that reads a
JSON formatted file and stores the data in a list data structure.

Find the name of the artist who produced the image for the card "Force
of Will" from the "Eternal Masters" set (three letter code: EMA).

*Hint*: Start of by loading the data into a variable (let's call it
 `cards`), examine the type of the variable, then use `names(cards)`
 to get a list of the names of the top level categories. Once you find
 the catrgory you need, use `class` and `names` to examine the
 sub-categories.

# Challenge 2: The Game of Life

John Conway developed the [Game of
Life](http://en.wikipedia.org/wiki/Conway%27s_Game_of_Life) to show
how complex systems can be obtained from a few simple rules.


The game has a grid of cells that are either alive or dead. At each
iteration of the game, the state of a cell is determined by the number
of neighbour cells that are alive. The rules are:

- If a cell has three living neighbours, it remains or becomes alive.
- If a cell has two living neighbours, it's state does not change.
- If a cell has less than 2 or more than three living neighbours, it dies.

The game is a zero player game, meaning that once we set up the
initial state are press go, we no longer interact with the game.

There are many initial states that end up with no life after a number
of iterations, but there are many interesting initial states that
grow, cycle or remain static. See the Wikipedia page for details of
these initial states.

We have written the code to run the game of life in R. To run it,
simply download the file [life.r](./life.r) into your *working
directory* and source the R script:
```{r eval=FALSE}
source("life.r")
```

Then click on the plot to show where you want the initial life to be,
then click Finish (at the top right corner of the plot). We then watch
the life grow and die.

Unfortunately, the script is missing the code from the function:
```{r eval=FALSE}
count.neighbours <- function(grid, pos) {
  ## Count the number of living neighbours of a point.
  
  counter = 0
  x = pos$x
  y = pos$y
  
  ### insert the missing code here!

  return(counter)
}
```
The function takes the life grid and a position. It returns the number
of living neighbours of the position. Note that each cell has eight
neighbours, so the number returned must be between 0 and 8.

Use your knowledge of R from the previous part of this lab to write
the missing code, then enjoy the game of life.  Make sure to try out
some of these [staring
patterns](http://en.wikipedia.org/wiki/Conway%27s_Game_of_Life#Examples_of_patterns).



