% Spatial Analysis of Twitter
% 300958 Social Web Analysis 
% Week 14 Lab Solutions

```{r setup, include=FALSE}
opts_chunk$set(fig.path='figure/week14-labsol-')
```

`r opts_chunk$set(prompt=TRUE, comment=NA, fig.cap="")`


- The file [flu.csv](flu.csv) contains the latitude and longitude of
800+ tweets containing the word "flu".  Using the `spatstat` package
construct an intensity map of the tweets and add a map of the world.
Try using different values for the bandwidth parameter `sigma`, for
example, sigma=1,5 and 10.

```{r}
library(maps, quietly = TRUE)
library(spatstat, quietly = TRUE)
flu.pts = read.csv("flu.csv")
flu.pp = ppp(flu.pts$lng, flu.pts$lat, 
             xrange=c(-180,180), yrange=c(-90,90))
den1 = density(flu.pp, sigma=5)
plot(den1, ribbon=FALSE, main="")
map('world', add=TRUE)
```


- The file [the.csv](the.csv) contains the latitude and longitude of
800+ tweets containing the word "the".  Construct a relative risk map
of the "flu"" tweets and "the" tweets, add a map of the world.  Try
using different values for the bandwidth parameter `sigma`.

```{r}
the.pts = read.csv("the.csv")
the.pp = ppp(the.pts$lng, the.pts$lat, 
             xrange=c(-180,180), yrange=c(-90,90))
both = superimpose(Athe=the.pp, Bflu=flu.pp)
r = relrisk(both, sigma=5)
plot(r, ribbon=FALSE, main="")
map('world', add=TRUE)
```

- The file [iphoneGEO.csv](iphoneGEO.csv) has the lat and long of the
  iphone 5 data. Try making an intensity plot of these tweets.

```{r}
iphone = read.csv("iphoneGEO.csv")
iphone.pp = ppp(iphone$lng, iphone$lat, 
             xrange=c(-180,180), yrange=c(-90,90))
plot(iphone.pp, cols=2)
map('world', add=TRUE)
den = density(iphone.pp, sigma=5)
plot(den, ribbon=FALSE, main="")
map('world', add=TRUE)
```
