% Assessing for Impact
% 300958 Social Web Analysis 
% Week 12 Lab

`r opts_chunk$set(prompt=TRUE, comment=NA, fig.cap="")`

# Searching using twitteR

Complete the usual OAuth regstration then conduct a twitter search for any string you like.
Try and pick a string that returns a largish number of results.

Using `twListToDF` convert the search results to a data frame. How many of your results are geocoded?

Plot the results on a map of the world, using the `maps` library. (see lectures)

# Searching for tweets around a specific location

Searching for tweets around a specific location requires us to construct a special string to code the location, 
using the latitude and longitude.

For example, using 

```{r include=FALSE,eval=FALSE}
require(twitteR)
load("../twitterOAuth.RData")
options(RCurlOptions = list(cainfo = system.file("CurlSSL", "cacert.pem",
                                                 package = "RCurl")))
registerTwitterOAuth(cred)
```

```{r eval=FALSE}
st = searchTwitter("Kevin Bacon", geocode= "-33.811352,151.025201,20km", n=1000)
```

searches for tweets  in a 20 kilometre radius around Parramatta.

Try this search for your string in the first part above. Did you get any hits?

# The Flu data.

The file [flu.csv](flu.csv) contains the latitude and longitude of 800+ tweets containing the word "flu".
Using the `spatstat` package construct an intensity map of the tweets and add a map of the world. 
Try using different values for the bandwidth parameter `sigma`, for example, sigma=1,5 and 10.

```{r include=FALSE, eval=FALSE}
require(maps)
require(spatstat)
flu.pts = read.csv("flu.csv")
flu.pp = ppp(flu.pts$lng, flu.pts$lat, 
             xrange=c(-180,180), yrange=c(-90,90))
den1 = density(flu.pp, sigma=5)
plot(den1, ribbon=FALSE, main="")
map('world', add=TRUE)
```


## Relative Risk

The file [the.csv](the.csv) contains the latitude and longitude of 800+ tweets containing the word "the".
Construct a relative risk map of the "flu"" tweets and "the" tweets, add a map of the world. 
Try using different values for the bandwidth parameter `sigma`.

```{r include=FALSE, eval=FALSE}
the.pts = read.csv("the.csv")
the.pp = ppp(the.pts$lng, the.pts$lat, 
             xrange=c(-180,180), yrange=c(-90,90))
both = superimpose(Athe=the.pp, Bflu=flu.pp)
r = relrisk(both, sigma=5)
plot(r, ribbon=FALSE, main="")
map('world', add=TRUE)
```

# Iphone5 data

The file [iphoneGEO.csv](iphoneGEO.csv) has the lat and long of the iphone 5 data. Try making an intensity plot of these tweets.

```{r include=FALSE, eval=FALSE}
iphone = read.csv("iphoneGEO.csv")
iphone.pp = ppp(iphone$lng, iphone$lat, 
             xrange=c(-180,180), yrange=c(-90,90))
plot(iphone.pp, cols=2)
map('world', add=TRUE)
den = density(iphone.pp, sigma=5)
plot(den, ribbon=FALSE, main="")
map('world', add=TRUE)
```
