% Twitter and Facebook APIs 
% 300958 Social Web Analysis 
% Week 4 Lab

`r opts_chunk$set(prompt=TRUE, comment=NA)`
`r knit_hooks$set(source = function(x, options) paste("<pre lang='r'>", x, "</pre>", sep = ""))`

# R Basics

## Installing R

All of the lab sessions for 300958 Social Web Analysis will use R and
RStudio. If these two applications are not installed on the computer
you are using, download them from:

- R: [http://cran.csiro.au/](http://cran.csiro.au/)
- RStudio: [http://www.rstudio.com/ide/download/desktop](http://www.rstudio.com/ide/download/desktop)

Install R, then Rstudio.

Once they are installed, open RStudio and say hello.
```{r}
print("Hello world!")
```

## Getting familiar with RStudio

Once RStudio is loaded, we will see four sections:

- Text editor (top left). Use this to enter code and save it as a script.
- Console (bottom left). Code can be entered straight into the console, or scripts can be run by entering source("script_filename.r")
- Workspace (top right). This section shows the variables that are currently defined. 
- Browser (bottom right). We can access files, view plots, examine packages and access help in the browser.

To write your first script, type:
```{r}
print("Hello world!")
```
into the text editor, save it as hello.r. Then go to the console section and type:
```{r}
source("hello.r")
```
By sourcing the script, you will run every command. Since there is one command in the script, only one command will be run.

If R complains that it cannot find the file, it may be looking in the wrong directory. To set the directory, use:
```{r}
setwd("directory")
```
where directory is where your script is stored.

## Writing functions

If a particular piece of code is run often, it may be wise to place it
in a function. In R, functions are stored in variables. If we define the function:
```{r}
addOne = function(x) {
       return(x + 1)
}
```
we can examine the function by typing its name in the console and pressing enter.
To run the function, we must add parentheses to the end and include any arguments:
```{r}
addOne(5)
```
If the result is not assigned to a variable, it is printed to the screen.


## Installing Libraries

Installing libraries requires access to the Internet and hence R needs
to know about the Web proxy when within UWS.

Information on how to tell R about the proxy is here:

- http://support.rstudio.org/help/kb/faq/configuring-r-to-use-an-http-proxy

Information about the proxy setting is here:

- http://www.uws.edu.au/information_technology_services/its/servicedesk/proxy_settings

Once this has been done, you may install libraries using the install.packages command:

install.packages("twitteR")

If R will not speak to the proxy, you can download the packages and install them locally.

Get the package from here:

- http://cran.r-project.org/web/packages/available_packages_by_name.html

Install the package by telling R where you saved it:

- install.packages("/tmp/bigdata_0.1.tgz", repos=NULL)


# Accessing Twitter

To access the Twitter API, a Twitter account is needed. If you do not have one, sign up here:

- https://twitter.com/

To access the Twitter API, we must register a program that is accessing the API. To register the program, visit this link:

- https://dev.twitter.com/apps

and log in. Once you have logged in, click "Create a new application"
and fill in the form. When the application is created, it will be
provided a consumer key and a consumer secret; these text sequences
are needed soon.

We can access the Twitter API in R by loading the twitteR and ROAuth libraries (and
installing them first if needed).

`r library("twitteR")`
```{r}
library("twitteR")
library("ROAuth")
```

We then need to set up the OAuth session details. Note: insert your
applications key and secret where needed.

```{.r}
cred <- OAuthFactory$new(consumerKey="insert here",
     consumerSecret="insert here",
     requestURL="https://api.twitter.com/oauth/request_token",
     accessURL="https://api.twitter.com/oauth/access_token",
     authURL="https://api.twitter.com/oauth/authorize")
cred$handshake()
```

Once the handshake is performed, R will ask you to visit a URL to get
an access code from Twitter. Get this code and provide it to R.
Once the handshake is done, we can save the credential information
and use it later, without having go directly to Twitter again.

```{.r}
save(cred, file="twitter.Rdata")
```

When we want to access Twitter in another session, we simply load and
register the credentials:
```{.r}
load("twitter.Rdata")
registerTwitterOAuth(cred)
```


```{.r}
x = searchTwitter('kevin bacon')

x = searchTwitter('social web analytics', n = 500, lang = "en")
z = twListToDF(x)



library("wordcloud")
library("tm")

cz = Corpus(VectorSource(z$text))
#cz = tm_map(cz, function(x) iconv(enc2utf8(x), sub = "byte"))
cz = tm_map(cz, function(x) iconv(x, to='UTF-8-MAC', sub='byte'))
cz <- tm_map(cz, removeNumbers)
cz <- tm_map(cz, removePunctuation)
cz <- tm_map(cz, stripWhitespace)
cz <- tm_map(cz, tolower)
cz <- tm_map(cz, removeWords, stopwords("english")) # this stopword file is at C:\Users\[username]\Documents\R\win-library\2.13\tm\stopwords 
cz <- tm_map(cz, removeWords, c("social", "web", "analytics"))
cz <- tm_map(cz, function(x) sub("http[^ ]*","", x))
cz <- tm_map(cz, stemDocument, language = "english")
wordcloud(cz)


library(tm) #load text mining library
setwd('F:/My Documents/My texts') #sets R's working directory to near where my files are
a  <-Corpus(DirSource("/My Documents/My texts"), readerControl = list(language="lat")) #specifies the exact folder where my text file(s) is for analysis with tm.
summary(a)  #check what went in
a <- tm_map(a, removeNumbers)
a <- tm_map(a, removePunctuation)
a <- tm_map(a , stripWhitespace)
a <- tm_map(a, tolower)
a <- tm_map(a, removeWords, stopwords("english")) # this stopword file is at C:\Users\[username]\Documents\R\win-library\2.13\tm\stopwords 
a <- tm_map(a, stemDocument, language = "english")
adtm <-DocumentTermMatrix(a) 
adtm <- removeSparseTerms(adtm, 0.75)


cz = Corpus(VectorSource(z$text))
tdm <- TermDocumentMatrix(cz, control = list(removePunctuation = TRUE, stopwords = TRUE, stem = TRUE))
tdm = weightTfIdf(tdm)
inspect(tdm)
tdm2 <- removeSparseTerms(tdm, sparse=0.95)
df <- as.data.frame(inspect(tdm2))
d = dist(df)
h = hclust(d)
rect.hclust(h, k=5, border="red")


cos.dissim <- function(x,y) {
	   x = as.numeric(x)
	   y = as.numeric(y)
	   return(1 - (x %*% y)/(sqrt(x %*% x)*sqrt(y %*% y)))
}

cos.sim <- function(x,y) {
	   x = as.numeric(x)
	   y = as.numeric(y)
	   return((x %*% y)/(sqrt(x %*% x)*sqrt(y %*% y)))
}

d = matrix(0, nrow(df), nrow(df))

for (a in 1:nrow(df)) {
    for (b in 1:nrow(df)) {
    	d[a,b] <- cos.dissim(df[a,],df[b,])
    }
}
rownames(d) <- rownames(df)

d = as.dist(d)
h = hclust(d)
plot(h)

```

http://files.meetup.com/1772780/Mining%20Twitter%20Using%20R.pdf

# Accessing Facebook

