% Introduction to R
% 300958 Social Web Analysis 
% Week 3 Lab Solutions

```{r setup, include=FALSE}
opts_chunk$set(fig.path='figure/week03-labsol-')
```

`r opts_chunk$set(prompt=TRUE, comment=NA)`
`r knit_hooks$set(source = function(x, options) paste("<pre lang='r'>", x, "</pre>", sep = ""))`


- What data structure would you choose to store your past results at
UWS? Write code to store your results in the variable `past.results`.

Out past results are tabular, so we would use a data frame.

```{r}
past.results = data.frame(
  UnitNumber = c(300700, 300958, 200045),
  UnitName = c("Statistical Decision Making", "Social Web Analytics", 
               "Quantitative Project"),
  Mark = c(73, 81, 48)
)
print(past.results)
```

- Write a for loop to convert a number to a sequence of plus signs. For
example, 5 will be converted to "+++++".

```{r}
n = 10
for (a in 1:n) {
  cat("+")
}
```


- Write an `if` statement to test if an item in `past.results` is a pass
or fail grade.  Then use a for loop to test all of the grades in
`past.results`.

```{r}
n = nrow(past.results)
for (a in 1:n) {
  if (past.results$Mark[a] >= 50) {
    cat(past.results$UnitNumber[a],":", as.character(past.results$UnitName[a]), 
        ":", "PASS", "\n")
  } else {
    cat(past.results$UnitNumber[a],":", as.character(past.results$UnitName[a]), 
        ":", "FAIL", "\n")
  }
}
```

- Write a function `check.pass` that takes the data structure
`past.results` and returns a vector showing if each grade was a pass
or a fail.

```{r}
check.pass = function(results) {
  n = nrow(results)
  passStatus = rep(0,n)
  for (a in 1:n) {
    if (results$Mark[a] >= 50) {
	  passStatus[a] = TRUE
    } else {
	  passStatus[a] = FALSE
	}
  }
  return(passStatus)
}
check.pass(past.results)
```


- Use your knowledge of R from the previous part of this lab to write
the missing code, then enjoy the game of life.  

```{r}
count.neighbours <- function(grid, pos) {
## Count the number of living neighbours of a point.

  counter = 0
  x = pos$x
  y = pos$y

  ## check the 3x3 grid centred at pos
  for (x in pos$x + c(-1,0,1)) {
    for (y in pos$y + c(-1,0,1)) {
      ## wrap the coordinates, in case the point is at the grid boundary
      wrap.pos = wrap.coordinates(grid, list(x = x, y = y))
      ## if there is life, add 1
      if (inspect(grid,wrap.pos) == 1) {
        counter = counter + 1
      }
    }
  }
  wrap.pos = wrap.coordinates(grid, pos)
  ## remove state of centre point
  counter = counter - inspect(grid,wrap.pos)
  return(counter)
}
```

