% Graphs 1: Graphs and their parameters
% 300958 Social Web Analysis 
% Week 6 Lab Solutions

```{r setup, include=FALSE}
opts_chunk$set(fig.path='figure/week06-labsol-')
```

- By visually examining the two graphs above, which looks denser? Use
the function `graph.density` to compute the density of each graph and
compare the results to your guess.

The ER graph looks to have higher density since it contains many more
edges.

```{r}
library("igraph", quietly=TRUE)
g.er = erdos.renyi.game(n = 100,p = 0.1)
g.ba = barabasi.game(n = 100, directed=FALSE)
graph.density(g.er)
graph.density(g.ba)
```

- The diameter is the longest shortest path. Which of the two graphs do
you expect to have the largest diameter?  Use the function `diameter`
to compute the diameter of each graph.

We expect the ER graph to have a smaller diameter because there are
many paths between each of the vertices.

```{r}
diameter(g.er)
diameter(g.ba)
```

- What do you expect the degree distribution of each graph to look like?
We can compute the degree of each vertex using the function `degree`.
We can also compute the degree distribution of the graph using the function
`degree.distribution`. Use the `help` pages to understand the output.

The ER degree distribution should look mound shaped (a mean with left
and right tails). The BA degree distribution should look exponentially
decaying (many vertices with low degree, a few with high degree).

```{r}
degree(g.er)
degree.distribution(g.er)
degree(g.ba)
degree.distribution(g.ba)
```

- Which vertex is most central according to Degree Centrality?

To find the most central, we order the vertices by their degree.
```{r}
order(degree(g.er), decreasing=TRUE)
order(degree(g.ba), decreasing=TRUE)
```


- Read the R help page for
`closeness` to find what R is computing. Then work out which of the
vertices is the most central with respect to closeness centrality.

The R function closeness provides the reciprocal of the sum of path lengths.
Therefore the sum of path lengths is:
```{r}
1/closeness(g.ba)
```
We want the vertex with the shortest path lengths, therefore we want
the maximum given by the R closeness function.
```{r}
order(closeness(g.er), decreasing = TRUE)
order(closeness(g.ba), decreasing = TRUE)
```


- Is the centre the same for all three centrality measures? Examine this for the 
Erdős-Renyi graph and Barabási–Albert graph.

```{r}
order(betweenness(g.er), decreasing = TRUE)
order(betweenness(g.ba), decreasing = TRUE)
```

Compare the above orders.



- Using the following graph:
```{r fig.cap="Small Graph"}
g3 = graph.formula(A-B, A-C, A-D, B-D, B-E, E-D, C-E)
plot(g3)
```

Calculate the Degree Distribution, Degree Centrality, Closeness
Centrality, Betweenness Centrality using the methods shown in the
lecture. Then check your answer using the R functions.


```{r}
degree.distribution(g3)
degree(g3)
1/closeness(g3)
betweenness(g3)
```


- Find the 10 friends that have the most followers. What are their
names?  Note the function `sort` will sort the vector of follower
counts. The function `order` will sort, but provide the position of
the sort. So to find the top 10, we use `order` with `decreasing=TRUE`
and choose the first ten values, giving us the positions of the top
10.

```{r eval=FALSE}
friendFollowCount = count.followers(friends)
friendPosition = order(friendFollowCount, decreasing = TRUE)[1:10]
topFriends = friends[friendPosition]
```

- Write a `for` loop to check if any of the friend accounts are
protected and store the `TRUE`/`FALSE` values in the variable `protected.status`. 

```{r eval=FALSE}
n = length(friends)
protected.status = rep(0, n)
for (a in 1:n) {
  protected.status = friends[[a]]$getProtected()
}
```

- Using the function `count.followers` find the 10 unprotected
  friends with the most followers and store them in `top.unprotected.friends`.

```{r eval=FALSE}
friendFollowCount = count.followers(unprotected.friends)
friendPosition = order(friendFollowCount, decreasing = TRUE)[1:10]
top.unprotected.friends = unprotected.friends[friendPosition]
```

- Write a `for` loop to download 100 friends from the 10 most
popular friends of Wil Wheaton and store them in `more.friends`.

```{r eval=FALSE}
n = length(top.unprotected.friends)
for (a in 1:n) {
  more.friends[[a]] = top.unprotected.friends[[a]]$getFriends(100)
}
```

- Write a `for` loop to store all 100 screen names in the variable `friend.names`.

```{r eval=FALSE}
friend.names = c()
n = length(friends)
for (a in 1:n) {
  friend.names[a] = friends[[a]]$getScreenName()
}
```


- Using what you have done above, write the function:

```{r eval=FALSE}
user.to.edgelist <- function(user, friends) {

  # create the list of friend screen names
  friend.names = c()		
  for (a in c(1:length(friends))) {
    friend.names[a] = friends[[a]]$getScreenName()
  } 

  user.name = rep(user$getScreenName(), length(friends)) # repeat user's name
  el = cbind(user.name,friend.names) # bind the columns to create a matrix

  return(el)
}
```

- Who is at the centre of the graph? Use the centrality measures to
  examine this.

```{r eval=FALSE}
order(closeness(g2))
```

- Examine the graph density. Is it sparse or dense?

```{r eval=FALSE}
density(g2)
```

- Examine the degree distribution. Is this graph more similar to an
Erdős-Renyi graph or a Barabási–Albert graph?

```{r eval=FALSE}
degree.distribution(g2)
```
