% Simple Exposure Analysis
% 300958 Social Web Analysis 
% Week 4 Lab

```{r setup, include=FALSE}
opts_chunk$set(fig.path='figure/week04-lab-')
```

`r opts_chunk$set(prompt=TRUE, comment=NA)`


# Using R
## Reading in Data

The easiest way to get data into R is to use `read.csv`. CSV stands for comma seperated values. 
By default, R will read CSV files and if a column is numeric, make a numeric variable. 
If the column contains any character data, the whole column will be made into a factor.

The Key Metrics sheet from facebook has a top row with short variable *names* and a 
second row with longer descriptions.
This means that R will read every column as a factor. To prevent this we use 
the `as.is=TRUE` option in `read.csv`. This has the effect of leaving character variables as character. 

This problem doesn't occur for most of the other sheets.

### Exercise
Use `read.csv` to read in the CSV files [keyMetrics.csv](keyMetrics.csv), [LifetimeLikesByGenderAge.csv](LifetimeLikesByGenderAge.csv) and [WeeklyReachDemog.csv](WeeklyReachDemog.csv) into data frames.
Make sure you give sensible names to the data frames you create. Use as.is=TRUE for keyMetrics.csv to avoid the mentioned problem.

## Manipulating data frames

The `dim` function in R gives the dimensions of a data.frame. Use it on your objects.
The `head` function shows the top few rows, and typing the data.frame name prints all of it out. Try these out.

The `summary` command gives a quick summary of each variable (column) in the data frame.

## Key metrics

To do anything with the Key Metrics data we have to convert the dates 
(they will have been read as character strings), and convert the other columns to numeric data.

### Converting Dates

The first column in Key Metrics is a date string. To use it in plots, it needs to be a data structure
that can be manipulated. R has several forms of date structures but we will use something called POSIXlt.
To convert from a string to POSIXlt we use the function `strptime`. Here is the code from lectures.

```{r eval=FALSE}
dates <- keyMetrics[,1]
dates <- dates[-1]
dates <- strptime(dates, format="%m/%d/%y")
```

We extract the first column, remove the first row (it is the long description) and convert using `strptime`.

Try this with your data.frame

### Extracting a column of numeric data

In lectures we extracted the Daily and Weekly Total Reach (columns 15 and 16) using the code like the following.

```{r eval=FALSE}
reach <- keyMetrics[,15]
reach <- as.numeric(reach[-1])
wreach <- as.numeric(keyMetrics[-1,16])
```

Identify and extract the "Weekly Total impressions" variable from the keyMetrics. HINT: the `names` function will
give a list of column names.

Impressions are the views of any content on a page.

### Plotting
Try plotting the "Weekly Total impressions" by date. Remember for reach we used `plot(dates, reach, type="l")`

```{r include=FALSE}
keyMetrics <- read.csv("keyMetrics.csv", as.is=TRUE)
dates <- strptime(keyMetrics[-1,1], format="%m/%d/%y")
impress <- as.numeric(keyMetrics[-1,28])
plot(dates,impress, type="l")
```

Look at the help page for `plot` (see the examples) 
and try changing the axis labels (`xlab=` and `ylab=`), give a title (`main=`) and
changing colours and line type (`col=` and `lty=`)

## Likes and Reach

The Likes and Reach by Demographics sheets (CSV files) are a different format. Generally we are 
interested in these demographics for a particular date.

The following template code will extract the 158th row, including only the females and males (columns 3 to 16).

```{r eval=FALSE}
tab <- matrix(as.numeric(WeekReach[158,3:16]), nrow=2, byrow=TRUE)
colnames(tab) <- c("13-17", "18-24", "25-34", "35-44", 
                   "45-54", "55-64", "65+")
rownames(tab) <- c("Female","Male")
print(tab)
```

Try this for your data frames, use different rows (between 1 and 158).

Use `barplot(tab, legend=TRUE, col=c("pink","lightblue"))` to make a barplot. 
Investigate changing the colours (`col=`) and the `beside=` option. 

# $\chi^2$ Test

In this section, we will use the $\chi^2$ test to identify properties
of the population data.

## Test for independence

We showed in the lectures that we can perform a test for independence
for a given two way table (two way meaning, has more than one row and
column). To perform the test, we need to:

- compute the expected values of the table if the rows and columns are
  independent (code shown in the lecture slides).
- shuffle the rows and columns of the table.

To peform the shuffling, we must first untabulate the table. For
example, if we start with the table:

|    | A | B | C |
|----|---|---|---|
| X  | 2 | 1 | 1 |
| Y  | 1 | 3 | 1 |

We must convert it to the form:

|Column | Row |
|:---:|:---:|
| A | X |
| A | X |
| A | Y |
| B | X |
| B | Y |
| B | Y |
| B | Y |
| C | X |
| C | Y |

Write the code to do this table conversion.

Hint: To compute the two column table, the two columns can be computed
seperately as `gender` and `age`, then combined using `tab2 =
data.frame(Gender = gender, Age = age)`. Also, the functions `rowSums`,
`colSums`, `rownames`, `colnames` and `rep` may be useful (if you are unfamiliar with these
functions, read the R documentation on them, e.g help(rowSums)).


Once we have the data in two columns, we shuffle the columns and recompute the table and compute the $\chi^2$ value (as shown in the lecture).

Using the above table `tab` and the hypotheses $H_0$: Gender and Age are independent, $H_A$: Gender and Age are not independent:

- Compute the $\chi^2$ randomisation distribution.
- Compute the $\chi^2$ statistic for `tab`.
- Compute the $p$ value of the test.

Finally state the conclusion of the test.




## Test for preference

We now want to test ($H_0$) if our site if preferred equally by all
age groups, or ($H_A$) is there a given age group that has a
difference proportion to the others.

We are only interested in age and not gender, so lets marginalise out the gender:
```{r eval = FALSE}
age = colSums(tab)
n = sum(age)
k = length(age)
```

To test if a given sample comes from a multinomial distribution, we use the $\chi^2$ test:
$$
	\chi^2 = {\sum_{i = 1}^{k} \frac{X_{i} - np_i}{np_i}}
$$
where $X_i$ are the frequencies in the sample, $n$ is the sample size, $k$ is the number of categories, and $p_i$ are the proportions assumed under $H_0$.

- Given the sample size `n` and the age table, compute the expected
  age group frequencies $np_i$ if all ages have the same proportion.

- Compute the $\chi^2$ value using the observed and expected frequencies.


We need to generate the randomisation distribution, given the
$H_0$. $H_0$ says that all proportions are equal, and we have a sample
of size `n`. Therefore, we can compute the randomisation distribution by:

- sample from the set of age categories with replacement using the given proportions $p_i$ and sample size $n$ (using `sample`).
- compute the $\chi^2$ value of the sample.
- Repeat at least 1000 times to obtain a distribution of $\chi^2$ values given $H_0$.


Once the distribution is obtained, compute the $p$ value by comparing
it to the sample $\chi^2$ value. Was the result as expected (does is
say that the ages group do not have equal proportions?).



# More R - a challenge

Construct the $2\times 2$ table containing the number of males and females
(independent of age), versus the months of April and May. Does a $\chi^2$ test
show that the reach for gender is independent of the two months?


