% Introduction to Social Networks
% Laurence A. F. Park and Glenn Stone
% March 22, 2005


# Algorithms in R

## Hello World

```{r}
Hello <- function() {
  print("Hello World")
}


Hello()
```

## R Scripts
In R commands or programs are stored in text files called "scripts".

They are usually given the file extension ".r" or ".R"

R itself has a menu item (under "File") to open a script or create a new one.

Rstudio also does.

## R Studio

R studio is an IDE (Integrated Development Environment)

\includegraphics[width=0.99\textwidth]{Rstudio.png}

## R Studio 
* The top left pane is an editor for scripts (and other files)
* The bottom left is the R console
* The top right has workspace tab and history tab
    * Workspace is the list of objects created in R
    * History contains the commands you have run
* The bottom left has 4 tabs
    * Files - a listing of the *current working directory*
    * Plots - any plots made in R go here
    * Packages - the available add-on packages
    * Help - help and documentation

## R Syntax - Basic data

R has several basic data types

* numbers - 1234, 3.14, 1e5
* strings - "Hello"
* factors - factor(c("H","H", "T", "H"), levels=c("H","T"))

## R Syntax - Operators

It has operators 

* Numeric,    + - * / ^
* Comparison,  < > <= >= == !=
* Logical,    & | && ||
* Assignment  <- = ->


### Problem

Given two real numbers $x$ and $y$, what is the R code to test if:
$$
	\log(x^2 + \pi) \ge y
$$

## R Syntax - Objects and Expressions

Objects have names made up of letters, numbers, . and _

A letter comes first, case matters

Values can be assigned to them

```{r}
x = 1
y <- "String"
```

An R expression is combination of objects, operators and data items (possibly with brackets)

```{r}
1+1
2*(1+x)
```

## R Syntax - Vectors

Vectors are set of items stored together. They have length and content.

They are constructed with "c"

```{r}
x <- c(1,3,7)
length(x)
```

The ":" operator can be used to set up a sequence of integers

```{r}
x <- 2:7
x
```

## R Syntax - Vectors

"seq" is more general (see help)

```{r}
x <- seq(2,12,2)
x
```

Individual entries are accessed with "[" (square brackets) and can be assigned to
```{r}
x[3]
x[4] <- 10
x
```

## R Syntax - Vectors

### Problem

How many ways can you think of creating the R vector:
$$
	[ ~ 0.1 ~ 0.2 ~ 0.3 ~ 0.4 ~ 0.5 ~ 0.6 ~ 0.7 ~ 0.8 ~ 0.9 ~ 1.0 ~ ]
$$


## R Syntax - Lists

Vectors contain things that are of the same type. Lists are more general.

```{r}
x <- list(1, "Hello World", 1:5)
x
```

## R Syntax - Lists

Entries are accessed (and assigned to) with "[["
```{r}
x[[2]]
x[[2]] <- "Good-bye!"
x
```

## R functions

R has Functions both built in and you can write your own.

Built in functions can be called on objects or vectors

```{r}
sqrt(2)
x <- 1:4
sqrt(x)
```
Note that many functions are *vectorized*

## R functions

Functions are called using brackets "()" (no space)

Any arguments are in the brackets, arguments can be named.

```{r}
sum(1:10)  ### Sum of 1 upto 10
sum(x=1:10)  ### Same
```

## R functions

To define a new function use the "function" keyword

```{r}
Hello <- function() {
      print("Hello World")
}
```
Note this function has no arguments. But we still need the brackets to call it
```{r}
Hello()
```

## R functions

### Problem

Write a function that takes two arguments $x$ and $y$ and returns a
vector, containing all of the numbers between $x$ and $y$, with a 0.1
interval (i.e. returns the set $z = x + 0.1n$ where $n$ is a positive
integer and $z \le y$).


## R Control structures

Control structures control how the function proceeds

### if statements
```{r, eval=FALSE}
if(condition) {
  expression~1
} else {
  expression~2
}
```
Where expression~1 is executed if the condition is TRUE or non-zero
and expression~2 is executed otherwise

## R Control structures
Loops allow things to be repeated. In R loops run through vectors.

### for loops
```{r, eval=FALSE}
for(i in vector) {
  expression
}
```
expression is executed with i set to each value in vector in turn

## Example 1 - Factorial

```{r, tidy=FALSE}
Factorial <- function(n) {
  fac <- 1
  for(i in 1:n) {
  	fac <- fac * i
  }
  return(fac)
}

Factorial(5)
```

## Example 2 - Recursive Factorial

```{r, tidy=FALSE}
RFactorial <- function(n) {
  if(n==1) {
    return(n)
  }
  else {
    return(n * RFactorial(n - 1))
  }
}

RFactorial(5)
```



## R Control structures
There are break and next statements:

* break: breaks out of a loop
* next: skips to the next entry

There are also while loops

### while loops
```{r, eval=FALSE}
while(condition) {
  expression
}
```
Keep executing the expression while the condition is TRUE (or nonzero)

## Towers of Hanoi example?


[Tower of Hanoi](https://en.wikipedia.org/wiki/Tower_of_Hanoi)

\includegraphics[width=0.99\textwidth]{HanoiStart.pdf}

## Towers of Hanoi example?
```{r, eval=FALSE, tidy=FALSE}
move.hanoi <- function(k, from, to, via) {
  if (k > 1) {
    move.hanoi(k - 1, from, via, to)
    move.hanoi(1, from, to, via)
    move.hanoi(k - 1, via, to, from)
  }
  else {
    cat("Move ", LETTERS[tower[[from]][1]], 
        " from ", from,
        " to ", to, "\n")
  }
}

move.hanoi(n, 1, 3, 2)
```

